---
title: "Exercises: Visualization 1"
output:
  learnr::tutorial:
    theme: flatly
    css: css/nowrap.css
runtime: shiny_prerendered
---

```{r base, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
tools::Rd2txt_options(underline_titles = FALSE, width = 75)
```

```{r setup, include=FALSE}
library(tidyverse)
tools::Rd2txt_options(underline_titles = FALSE, width = 75)
```

## Introduction

These exercises introduce visualisation with *ggplot2*, which is part of the *tidyverse*. In many cases a part of the code is given, and your task is to replace the underscores `___` or the dots `...` with something meaningful before you can run the code. All exercises are self-contained.

Each code snippet can be run by clicking the green "Run Code" button or pressing Ctrl + Enter. In some cases the exercises contain hints or the solution. There's no automatic validation for the R code. If you see a red error message containing "unexpected input", this means that you need to replace one or more of the `___` placeholders.

The *tidyverse* package is loaded for all exercises:

```r
library(tidyverse)
```

Try it out!

```{r hello-world,  exercise = TRUE, exercise.lines = 8}
print("Hello, world!")
```

## Basic arithmetics

Use R as a calculator. Evaluate a few arithmetic expressions.

```{r calc, exercise = TRUE}
1 + 2
```

Do operator precendence rules hold? How do you override the default precedence?

```{r precedence, exercise = TRUE}
___ + ___ * ___
___
```

```{r precedence-hint}
(___ + ___) * ___
```

Also try the modulo operator `%%`, the integer division `%/%`, and functions such as `floor()`, `sin()`, `pmax()`.

```{r custom-ops, exercise = TRUE}
```

Explain what happens when you run the following two examples:

```{r nl-after-plus, exercise = TRUE}
5 +
  3
```

```{r nl-before-plus, exercise = TRUE}
5
+ 3
```

## Getting help

Look up the help for the following objects and functions:

- `diamonds`
- `geom_boxplot()`
- `near()`
- `%%` (use backticks or quotes)

```{r help, exercise = TRUE}
?diamonds
```

Run the example for `near()` by pasting it into the code window below:

```{r example-manual, exercise = TRUE}

```

Find out what the `example()` function does. Can you use it to run the example for `near()`?

```{r example-automatic, exercise = TRUE}

```

Start typing `geom_` and see the functions available. (You may need to press Ctrl + Space.) Show the help for some of these functions:

```{r example-autocomplete, exercise = TRUE}

```

Visit https://rdocumentation.org, search for functions and objects from this exercise.

Use the search term `geom_boxplot site:tidyverse.org` in a search engine.

## Scatterplots

Show the `mpg` dataset.

```{r mpg, exercise = TRUE}
mpg
```

Create a few scatterplots using the following template (here with `hwy` versus `displ`). Share particularly interesting or boring examples you might encounter.  Did you find anything surprising?

Hint: Look up the contents of the columns in the help.

```{r scatter, exercise = TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

Plot highway fuel economy given as liters per 100 kilometers against engine displacement.

Hint: Use the formula `235 / hwy` to convert from miles per gallon.

```{r scatter-liters, exercise = TRUE}
ggplot(data = ___) +
  geom_point(mapping = aes(x = displ, y = ___))
```

Find more examples in [Section 3.2.4](http://r4ds.had.co.nz/data-visualisation.html#exercises) of the "R for data science" book.

```{r scatter-r4ds, exercise = TRUE, exercise.lines = 10}
```

## 3D+ Scatterplots

In the `hwy` vs. `displ` plot, map an additional variable to the "color" aesthetic.  Which cars consume more fuel than expected by the general trend?

```{r scatter-color, exercise = TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = ___))
```

Experiment with the "color", "shape", "size", and "alpha" aesthetics.  Which combinations of attribute class (categorical/continuous) and aesthetics work well, which don't? Expand on the more surprising examples in the previous exercises.

Hint: Use `factor(___)` to convert continuous variables with a limited set of values to categorical variables.

```{r scatter-color-more, exercise = TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(___))
```

Can you change both color and shape at the same time? What about the other aesthetics?

```{r scatter-color-shape, exercise = TRUE}
ggplot(___) +
  geom_point(___)
```

What happens if you map the same variable to more than one aesthetic?

```{r scatter-color-repeat, exercise = TRUE}
ggplot(___) +
  geom_point(___)
```

## Manual aesthetics

Plot `hwy` vs. `displ` with approx. 1/3 opacity for each point, in blue.

Hint: Use quotes `"` for the color.

```{r scatter-manual, exercise = TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    color = "___",
    alpha = ___
  )
```

What happens if you assign a variable, e.g. `year`, to an aesthetic *outside* the `aes()` call?

```{r scatter-manual-outside, exercise = TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(___),
    color = year
  )
```

What happens if you assign a constant, e.g. `3`, to the "size" aesthetic *inside* the `aes()` call?

```{r scatter-manual-inside, exercise = TRUE}
ggplot(___) +
  geom_point(
    mapping = aes(___, size = 3)
  )
```

What values are valid for `color`, `alpha`, `shape` and `size`?

```{r scatter-manual-valid, exercise = TRUE}
ggplot(___) +
  geom_point(
    mapping = ___,
    ___
  )
```

Find more exercises in [Section 3.3.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-1) of r4ds.

```{r scatter-manual-r4ds, exercise = TRUE, exercise.lines = 10}
```

## Other point geoms

Try `geom_smooth()`. What does it do?

```{r point-smooth, exercise = TRUE}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

What do the arguments `se` and `method` to `geom_smooth()` change?

```{r point-se, exercise = TRUE}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy),
    se = ___,
    method = ___
  )
```

Can you display both the points and the smoothing line?

```{r point-smooth-points, exercise = TRUE}
ggplot(data = mpg) +
  geom_point(___) +
  geom_smooth(___)
```

How does the order of the `geom_...()` calls affect the display?

```{r point-order, exercise = TRUE}
ggplot(data = mpg) +
  ___ +
  ___
```

What does `geom_rug()` do?

```{r point-rug, exercise = TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_rug(mapping = ___)
```

Try specifying the mapping in the `ggplot()` call. What's the difference?

```{r point-rug-one, exercise = TRUE}
ggplot(data = ___, mapping = ___) +
  geom_point() +
  geom_rug()
```

Try to reduce overplotting with transparency. How do you draw jittered points?

```{r point-rug-overplot, exercise = TRUE}
ggplot(___, ___) +
  ___() +
  geom_rug(___)
```

Can you plot both highway and city economy against displacement in one plot?

Hint: The solution to this exercise is *not* the recommended way of doing this in ggplot2.  We'll find a better way in a subsequent exercise.

```{r point-rug-hwy-cty, exercise = TRUE}
___ +
  ___(___) +
  ___(___)
```

Find more exercises in [Section 3.6.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-3) of r4ds.

```{r point-r4ds, exercise = TRUE, exercise.lines = 10}
```


## Statistical transformations

Use a bar plot to find out how many cars of each drivetrain (front/rear/4wd) the `mpg` dataset contains. Which aesthetic mapping(s) do you need to specify? 

```{r stat, exercise = TRUE}
ggplot(data = mpg, mapping = aes(___)) +
  geom_bar()
```

1. Does the appearance of the plot change when you add `y = ..count..` to the `aes()` call? Why/why not?

```{r stat-count, exercise = TRUE}
ggplot(___, ___) +
  geom_bar()
```

1. What happens if you instead use `y = ..prop.., group = 1` in the `aes()` call? What happens if you omit `group = 1`? Why?

Hint: The section "Computed variables" in the help to `geom_bar()` offers a brief explanation.

```{r stat-prop, exercise = TRUE}
```

1. Visualize the overall distribution of fuel economy in the dataset with a histogram.  Compare this with a frequency polygon (`geom_freqpoly()`) in a second layer.

```{r stat-hist, exercise = TRUE}
___(___, mapping = aes(x = hwy)) +
  ___() +
  ___()
```

1. Visualize the distribution of fuel economy in the dataset per drivetrain.  Do you prefer a histogram or a frequency polygon? Why?

```{r stat-per-drv, exercise = TRUE}
___(___, mapping = aes(x = hwy, ___)) +
  ___()
```

Find more exercises in [Section 3.7.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-4) of r4ds.

```{r stat-r4ds, exercise = TRUE, exercise.lines = 10}
```

## Position adjustments

What's the most prevalent number of gears for manual or automatic transmissions?

```{r pos-gears, exercise = TRUE}
ggplot(data = mpg, mapping = aes(___)) +
  geom_bar()
```

```{r pos-gears-hint, eval = FALSE}
aes(x = trans)
```


Which aesthetic can you map to further discriminate by car class? Which position adjustment is most useful to detect missing combinations of drivetrain and car class? Try `"stack"`, `"dodge"` and `"fill"`, and explain the difference.

```{r pos-gears-by-class, exercise = TRUE}
ggplot(data = mpg, mapping = aes(___, ___)) +
  geom_bar(position = "___")
```

Draw a boxplot of highway fuel economy versus drivetrain.  Is fuel economy also affected by the number of cylinders? What is the default position adjustment for the boxplot geometry? What happens if you omit `factor()`?

```{r pos-hwy-by-drivetrain, exercise = TRUE}
ggplot(___, mapping = aes(x = drv, ___, fill = factor(___))) +
  ___()
```

Find more exercises in [Section 3.8.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-5) of r4ds.

```{r pos-r4ds, exercise = TRUE, exercise.lines = 10}
```

## Faceting

Has fuel economy changed considerably between 1999 and 2008?

```{r facet-years, exercise = TRUE}
ggplot(___, ___) +
  geom_histogram() +
  facet_wrap(~year)
```

Perhaps there is a difference if you also consider the car class?

```{r facet-years-class, exercise = TRUE}
ggplot(___, ___) +
  geom_histogram() +
  facet_grid(year ~ class)
```

Experiment with `facet_wrap()`, `facet_grid()`, aesthetic mappings, and smoothing layers.

```{r facet-years-class-experiment, exercise = TRUE}
ggplot(___, ___) +
  ___() +
  ___()
```

What changes if you add the argument `labeller = "label_both"` to the `facet_wrap()` call?

```{r facet-years-labeller, exercise = TRUE}
___ +
  ___() +
  facet_wrap(~year, ___)
```

Is it possible to use a different x and/or y scale for each facet? How?

```{r facet-years-free, exercise = TRUE}
___ +
  ___() +
  facet_wrap(~year, ___)
```

Experiment with other arguments to `facet_wrap()` and `facet_grid()`.

```{r facet-experiment, exercise = TRUE, exercise.lines = 10}
```

Find more exercises in [Section 3.5.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-2) of r4ds.

```{r facet-r4ds, exercise = TRUE, exercise.lines = 10}
```


## Scales

Give the scales in this plot a proper name.

```{r scales-simple, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  scale_...(___) +
  scale_...(___) +
  scale_...(___)
```

What happens if you call the same `scale_...()` function twice in a plot?

```{r scales-twice, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  scale_...(___) +
  scale_...(___)
```

Choose one of the available Brewer scales below for the color:

```{r}
RColorBrewer::display.brewer.all(colorblindFriendly = TRUE)
```


```{r scales-brewer, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  scale_..._brewer(palette = "___")
```

## Coordinate systems

Create a flipped boxplot of fuel economy by transmission type.

```{r coord-simple, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = ___, y = ___)) +
  geom_boxplot() +
  coord_...()
```

Plot highway vs. city fuel economy with an additional `geom_abline()` layer.  How do you set a 45° degree for the diagonal line? 

```{r coord-ab, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = ___, y = ___)) +
  geom_point() +
  geom_abline(color = "red") +
  coord_...()
```

Try to "estimate" an approximate ratio between the two economy measures by tweaking the `slope` argument to `geom_abline()`.

```{r coord-ab-estimate, exercise = TRUE}
___ +
  ___() +
  geom_abline(color = "red", slope = ___) +
  ___()
```

## Theming

Choose your favorite among the predefined themes.

Hint: All start with `theme_...()`, but watch out for `theme_set()`.

```{r theme-test, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  theme_...()
```

Apply `theme(legend.position = "bottom")` to move the legend to the bottom.

```{r theme-bottom, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  theme(___)
```

What happens if you then apply `theme_bw()`? Why?

```{r theme-bottom-bw, exercise = TRUE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  theme(___) +
  theme_bw()
```
