```{r 14-remove-all, include = FALSE}
rm(list = ls())
```


## One variable plots

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### One variable plots

```
```{r 14-setup-14-one-var-r}
library(tidyverse)
```
</details>

This type of plots visualizes ONE variable in a certain way.

To do this in a 2D space, a **statistical transformation** of the variable is required for the missing axis.

### Continuous variables

- Histogram: Most common way - grouping the variable into equal bins

- `geom_density()`, `geom_freq()`, `geom_dotplot()` and `geom_area()` are mainly doing the same as `geom_hist()`

We supply only _one_ variable to the `mapping` argument with the help of `aes()`.
This one is automatically grouped into 30 bins.

```{r include = FALSE}
# Continuous variables: histogram most common
```
```{r 14-continuous-variables-histogram-most-common}
ggplot(mpg, aes(x = hwy)) +
  geom_histogram()
```

---

```{r include = FALSE}
# showcase: density
```
```{r 14-showcase-density}
ggplot(mpg, aes(x = hwy)) +
  geom_density()
```

### Discrete variables

For discrete data, there is actually only one visualization method - the bar plot.

  \smallskip

*Note the difference of `geom_bar()` compared to `geom_hist()`.*

```{r include = FALSE}
# example for discrete var: barplot
```
```{r 14-example-for-discrete-var-barplot}
ggplot(mpg, aes(fl)) +
  geom_bar()
```

**Exercises**

```{r include = FALSE}
# Exercise
```
```{r 14-exercise}
browseURL("https://krlmlr.github.io/vistransrep/2019-11-zhr/scatter.html")
```
