```{r 19-remove-all, include = FALSE}
rm(list = ls())
```


## Export & saving

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Export & saving

```
```{r 19-setup-19-export-r}
library(tidyverse)
```
</details>


The default way to export plots in {{ggplot2}} is by using `ggsave()`. 

It differs slightly from other "exporting" functions in R because it comes with some smart defaults:

> ggsave() is a convenient function for saving a plot. It defaults to **saving the last plot** that you displayed, using the size of the current graphics device. It also **guesses the type** of graphics device from the extension.

```{r include = FALSE}
## create plot for export
```
```{r 19-create-plot-for-export}
ggplot(mtcars, aes(mpg, wt)) +
  geom_point()
```


```{r include = FALSE}
# show exporting
```
```{r 19-show-exporting}
ggsave("mtcars.pdf")
ggsave("mtcars.png")
```

```{r echo = FALSE, message=FALSE, results='hide'}
file.remove(c("mtcars.pdf", "mtcars.png"))
```

This might seem natural to you but is is not.
Let's compare base R and {{ggplot2}}.

### Base R vs. {{ggplot2}}

In base R

- one needs to open a specific graphic device first

- then create the plot 

- and close the graphic device again.

```{r include = FALSE}
# show how base plot exporting works
```
```{r 19-show-how-base-plot-exporting-works, eval = FALSE, out.width="100%"}
png("Plot.png")
plot(mpg$displ, mpg$hwy)
dev.off()
```


```{r include = FALSE}
# show how ggplot exporting works
```
```{r 19-ggplot-mpg-aes-disply-hwy, eval = FALSE}
ggplot(mpg, aes(disply, hwy)) +
  geom_point()
ggsave("Plot.png")
```

Base R plotting functions come with suboptimal defaults

- saving in pixels (differs on every monitors)

- saving as a square image

- no option to specify the DPI (dots per inch)

### Storing the plot as an R object

One of the major advantages of `ggplot()` is that you can save a plot as an R object and modify it later.

This is not possible with base R plots.


```{r include = FALSE}
# ggplot plots can be stored in R objects
```
```{r 19-ggplot-plots-can-be-stored-in-r-objects}
p <- ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```


```{r include = FALSE}
# saved ggplot objects can be used later on
```
```{r 19-saved-ggplot-objects-can-be-used-later-on}
p + geom_point(aes(color = class))
```


```{r include = FALSE}
# print the ggplot obj
```
```{r 19-print-the-ggplot-obj}
print(p)
```


```{r include = FALSE}
# look at ggplot2 obj structure
```
```{r 19-look-at-ggplot-2-obj-structure}
str(p)
```

### Best practices for exporting

Some best practices:

- Use a reasonable high DPI. A value of "300" is ok in most cases.

- Save in "inches" and not in "pixels". The latter always differs on screens with different resolutions (`png()` uses pixels by default.)

- Always specify a file name to ensure the right plot is chosen. Do not rely on the default behavior of `ggsave()` (even though it might seem convenient) which takes the last visualized plot.

- An alternative to `ggsave()` is `cowplot::save_plot()` which comes with sensible defaults for multi-plot arrangements.
