```{r 11-remove-all, include = FALSE}
rm(list = ls())
```


## Basics for visualisation in R using {ggplot2}

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Basics for visualisation in R using {ggplot2}

```
```{r 11-setup-11-ggplot-skeleton-r}
library(tidyverse)
```
</details>

<br>

In the {tidyverse} the standard package for visualization is {ggplot2}.
The functions of this package follow a quite unique logic (the "grammar of graphics") and therefore require a special syntax.
In this section we want to give a short introduction, how to get started with {ggplot2}.

```{r include = FALSE}
# # One time:
# install.packages("tidyverse") OR install.packages("ggplot2")
```
```{r 11-one-time}

```

### Creating the plot skeleton: `ggplot()`

The main function in the package is `ggplot()`, which prepares/creates a graph.
By setting the arguments of the function, you can:

1. Choose the dataset to be plotted (argument `data`)
2. Choose the mapping of the variables to the axes (or further forms of setting apart data) in the argument `mapping`. 
  This argument takes the result of the function `aes()`, which you will get to know in many different examples.


```{r include = FALSE}
# Empty graph
```
```{r 11-empty-graph}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
)
```

This created only an empty plot, because we did not tell {ggplot2} which geometry we want to use to display the variables we set in the `ggplot()` call.
We do this by adding (with the help of the `+` operator after the `ggplot()`-call) a different function starting with `geom_` to provide this information.

```{r include = FALSE}
# visualize variables as points: scatterplot
```
```{r 11-visualize-variables-as-points-scatterplot}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point()
```

This is maybe the most basic plot you can create.
To map a different variable than `disp` to the x-axis, change the respective variable name in the `aes()` argument.

```{r include = FALSE}
# aes() determines which variables are plotted
```
```{r 11-aes-determines-which-variables-are-plotted}
ggplot(
  data = mpg,
  mapping = aes(x = cyl, y = hwy)
) +
  geom_point()
```

You can exchange the variables to be plotted freely, without changing anything else to the rest of the code.

```{r include = FALSE}
# Pattern unchanged if only variables are plotted
```
```{r 11-pattern-unchanged-if-only-variables-are-plotted}
ggplot(
  data = mpg,
  mapping = aes(x = hwy, y = cty)
) +
  geom_point()
```

Always good to have: The *ggplot2* cheatsheet (https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf).

### What is a "statistical graphic"?

Wilkinson (2005) defines a grammar to describe the basic elements of a
statistical graphic:

> "[...] a statistical graphic is a mapping from data to
> aesthetic attributes (colour, shape, size) of geometric objects
> (points, line, bars)."

\hfill (Wickham, 2009)

### Terminology

- **Data:** The data to visualize -- consists of variables and observations.

- **Geoms:** Geometric objects which represent the data (points, lines, polygons, etc.).

- **Mappings:** Match variables with aesthetic attributes of the (geometric) objects.

- **Scales:** Mapping of the "data units" to "physical units" of the geometric objects (e.g. length, diameter or color); defines the _legend_.

- **Coord:** System of coordinates, mapping of the data to a two dimensional plain of the graphic; defines the _axes_ and _grid_.

- **Stats:** Statistical transformation of the data (5 point summary, classification, etc.).

- **Facetting:** Division and illustration of data subsets, also known as "Trellis" images.

### The Grammar of graphics ...

**is ...**

a formal guideline which describes the dependencies between all elements of a
  statistical graphic.

<br>

**isn't ...**

- a manual which tells us _which graphic_ should be created for a given question.
- a specification _how_ a statistical graphic should look like.

### About {ggplot2}

```{r, eval=FALSE, echo = FALSE}
library("ggplot2")
packageDescription("ggplot2")
```


```{r, eval = T, echo=FALSE, size="tiny", R.options=list(width=90)}
library("ggplot2")
p <- packageDescription("ggplot2")
p$Collate <- p$LazyLoad <- p$LazyData <- p$Built <-
    p$Suggests <- p$Extends <- p$Packaged <- p$Repository <-
    p$Description <- p$NeedsCompilation <- p$VignetteBuilder <- NULL
p[["Date/Publication"]] <- NULL
p[["Authors@R"]] <- NULL
p[["Enhances"]] <- NULL
p[["RoxygenNote"]] <- NULL
attr(p, "file") <- NULL
p
```
