```{r 11-remove-all, include = FALSE}
rm(list = ls())
```


## Basics for visualisation in R using {ggplot2}

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Basics for visualisation in R using {ggplot2}

```
```{r 11-setup-11-mpg-r}
library(tidyverse)
```
</details>

In the {tidyverse} the standard package for visualisation is {ggplot2}.
The functions of this package follow a quite unique logic (the "grammar of graphics") and therefore require a special syntax.
In this section we want to give a short introduction, how to get started with {ggplot2}.

```{r include = FALSE}
# # One time:
# install.packages("tidyverse")
```
```{r 11-one-time}

```

The main function in the package is `ggplot()`, which prepares/creates a graph.
By setting the parameters of the function, you can:
1. choose the dataset to be plotted in the parameter `data`
2. choose the mapping of the variables to the axes (or further forms of setting apart data) in the parameter `mapping`. This parameter takes the result of the function `aes()`, which you will get to know in many different examples.

```{r include = FALSE}
# Empty graph
```
```{r 11-empty-graph}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
)
```

This created only an empty plot, because we did not tell {ggplot2}, which geometry we want to use to display the result.
We do this by adding (literally using `+` after the `ggplot()`-call) a different function starting with `geom_` to provide this information.

```{r include = FALSE}
# Scatterplot
```
```{r 11-scatterplot}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point()
```

This is maybe the most basic plot you can create.
To map a different variable than `disp` to the x-axis, change the respective parameter in `aes()`.

```{r include = FALSE}
# aes() determines which variables are plotted
```
```{r 11-aes-determines-which-variables-are-plotted}
ggplot(
  data = mpg,
  mapping = aes(x = cyl, y = hwy)
) +
  geom_point()
```

You can exchange the variables to be plotted freely, without changing anything else to the rest of the code.

```{r include = FALSE}
# Pattern unchanged if only variables are plotted
```
```{r 11-pattern-unchanged-if-only-variables-are-plotted}
ggplot(
  data = mpg,
  mapping = aes(x = hwy, y = cty)
) +
  geom_point()
```
