```{r 12-remove-all, include = FALSE}
rm(list = ls())
```


## `geom_*` functions

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### `geom_*` functions

```
```{r 12-setup-12-geoms-r}
library(tidyverse)
```
</details>

`geom_*` functions are added to the main `ggplot()` call via the "+" operator and
(usually) placed on a new line.

A list of all available "geoms" can be found here:

https://ggplot2.tidyverse.org/reference/#section-layer-geoms

The most popular ones are

- `geom_point()`

- `geom_histogram()`

- `geom_boxplot()`

- `geom_bar()`

---

The `geom_*` family can be divided into three parts:

**One variable plots**

- `geom_hist()`
- `geom_bar()`
- etc.

**Two variable plots**

- `geom_point()`
- `geom_line()`
- `geom_boxplot()`
- etc.

**Three variables plots**

- `geom_raster()`
- `geom_sf()`
- `geom_tile()`
- etc.

### Arguments

```{r, eval=FALSE}
ggplot(data, mapping = aes(), ...) +
  geom_XXX(mapping = NULL, data = NULL, stat, ...)
```
```{r 12-ggplot-data-mapping-aes}

```

`geom_*` functions have the same basic arguments as `ggplot()`.
In addition, they come with more arguments specific to the respective "geom".

**stat**

The `stat` parameter defines a statistical transformation: 

- if set to `"identity"`: No transformation

- if set to `boxplot`: Boxplot transformation

- etc.

**position**

The same applies to the `position` argument. 
In the example below, points are not adjusted and just visualized where they appear in the data.

In the case of boxplots, a special position arrangement function is used to arrange everything nicely: `position_dodge2()` (here denoted by `position = "dodge2"`).

```{r eval = FALSE}
geom_point(mapping = NULL, data = NULL, stat = "identity",
  position = "identity", ..., na.rm = FALSE, show.legend = NA,
  inherit.aes = TRUE)

geom_boxplot(mapping = NULL, data = NULL, stat = "boxplot",
  position = "dodge2", ..., outlier.colour = NULL,
  outlier.color = NULL, outlier.fill = NULL, outlier.shape = 19,
  outlier.size = 1.5, outlier.stroke = 0.5, outlier.alpha = NULL,
  notch = FALSE, notchwidth = 0.5, varwidth = FALSE, na.rm = FALSE,
  show.legend = NA, inherit.aes = TRUE)
```
---

`geom_boxplot()` needs one variable to be of class `character` or `factor` (better) to initiate the grouping.


```{r include = FALSE}
# geom_boxplot needs a factor var
```
```{r 12-geom-boxplot-needs-a-factor-var}
class(mpg$class)
```


```{r include = FALSE}
# show geom_boxplot
```
```{r 12-show-geom-boxplot}
ggplot(mpg, aes(x = class, y = displ)) +
  geom_boxplot()
```

### Combining geoms

Multiple `geom_*` functions can be used in one plot.
A combination that is often used together is `geom_point()` and `geom_smooth()`

```{r include = FALSE}
# combining multiple geoms
```
```{r 12-combining-multiple-geoms}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Unless specified differently in the `geom_*()` call, all geoms will use the same variables.

### Summary

The modular principle of `ggplot2` enables:

- the combination of any geometric objects (geoms).
- a high flexibility and customizability

An extensive description of all geometric objects can be found on the `ggplot2` website https://ggplot2.tidyverse.org/reference/.

**Exercises**

https://krlmlr.github.io/vistransrep/2019-11-zhr/geoms.html
