```{r 14-remove-all, include = FALSE}
rm(list = ls())
```


## Statistical summaries

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Statistical summaries

```
```{r 14-setup-14-stat-r}
library(tidyverse)
```
</details>

Some `geom()`-functions do some transformations and calculations behind the scenes before the data are visualized.
One such function is `geom_smooth()` from the last section. 
Also `geom_bar()` does some transformations first.
It just needs a mapping for the x-axis.
Once that is done, the y-values displayed for each value of the x-variable are the numbers of of times that the x-value occurs in the dataset.
So basically the different values are counted.

```{r include = FALSE}
# geom_bar() for counting unique values
```
```{r 14-geom-bar-for-counting-unique-values}
ggplot(
  data = mpg,
  mapping = aes(x = class)
) +
  geom_bar()
```


```{r include = FALSE}
# geom_histogram() for counting values in bins
```
```{r 14-geom-histogram-for-counting-values-in-bins}
ggplot(
  data = mpg,
  mapping = aes(x = hwy)
) +
  geom_histogram(bins = 20)
```


```{r include = FALSE}
# geom_boxplot() for another way of summarizing numeric values
```
```{r 14-geom-boxplot-for-another-way-of-summarizing-numeric-values}
ggplot(
  data = mpg,
  mapping = aes(y = hwy)
) +
  geom_boxplot()
```


```{r include = FALSE}
# x-axis in geom_boxplot() to go into detail
```
```{r 14-x-axis-in-geom-boxplot-to-go-into-detail}
ggplot(
  data = mpg,
  mapping = aes(x = class, y = hwy)
) +
  geom_boxplot()
```


```{r include = FALSE}
# geom_smooth() for drawing a smooth curve through scatter plot
```
```{r 14-geom-smooth-for-drawing-a-smooth-curve-through-scatter-plot}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  geom_smooth()
```
