```{r 15-remove-all, include = FALSE}
rm(list = ls())
```


## Facet plots

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Facet plots

```
```{r 15-setup-15-facet-r}
library(tidyverse)
```
</details>

One helpful feature that {ggplot2} offers is the facet plot.
This means, that for each value of a given variable a different sub-plot is created.
The relevant functions for this are `facet_wrap()` and `facet_grid()`.
The former can be used like this.

```{r include = FALSE}
# facet_wrap()
```
```{r 15-facet-wrap}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  facet_wrap(~ class)
```

The latter has a few more options that you can play with.
Primarily, you can pass two variables to the function, one of which will determine the facetting along the rows and the other will determine the column facetting.

```{r include = FALSE}
# facet_grid()
```
```{r 15-facet-grid}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  facet_grid(year ~ drv)
```

The first argument of `facet_grid()` is by default `rows`.
If you pass a variable (needs to be the result of a call to `vars()` with the unquoted variable name as its parameter) to it, as many plots will be horizontally lined up, as the variable has distinct values.

```{r include = FALSE}
# facet_grid() only set rows
```
```{r 15-facet-grid-only-set-rows}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  facet_grid(vars(year))
```

Specifically choose parameter `cols` of `facet_grid()` to vertically stack the plots.

```{r include = FALSE}
# facet_grid() only set cols
```
```{r 15-facet-grid-only-set-cols}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  facet_grid(cols = vars(year))
```
