```{r 15-remove-all, include = FALSE}
rm(list = ls())
```


## Facet plots

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Facet plots

```
```{r 15-setup-15-facet-r}
library(tidyverse)
```
</details>

One helpful feature that {ggplot2} offers is the facet plot.
This means that for each value of a given variable a different sub-plot is created.
The relevant functions for this are `facet_wrap()` and `facet_grid()`.
The former aligns all sub-plots from left to right and from top to bottom by default.

```{r include = FALSE}
# facet_wrap()
```
```{r 15-facet-wrap}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  facet_wrap(vars(class))
```

Combinations of variables are also supported, and result in a sub-plot for each unique combination with valid data points.

```{r include = FALSE}
# facet_wrap() with multiple variables
```
```{r 15-facet-wrap-with-multiple-variables}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  facet_wrap(vars(class, drv))
```


The latter arranges all sub-plots in a grid and requires two variables.

```{r include = FALSE}
# facet_grid()
```
```{r 15-facet-grid}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  facet_grid(vars(year), vars(drv))
```
