```{r 16-remove-all, include = FALSE}
rm(list = ls())
```


## Labels

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Labels

```
```{r 16-setup-16-labels-r}
library(tidyverse)
```
</details>

For character variables there is further way of integrating its value to a plot.
`geom_text()` takes a `label` argument, which influences the plot in the following way.

```{r include = FALSE}
# showcase geom_text()
```
```{r 16-showcase-geom-text}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_text(label = "A")
```

Let's try to map this argument to a variable (here: `drv`) of our dataset in the `mapping` argument of `ggplot()`.

```{r include = FALSE}
# object vs string again
```
```{r 16-object-vs-string-again}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_text(label = "drv")
```

Right, of course we need to pass the variable without quotation marks, otherwise it is interpreted as a (constant) character variable.
When changing this, a vector with the values of the variable is passed on to `geom_text()`.
This is one way of including the values of character variables in a plot.

```{r include = FALSE}
# dynamic labeling to variable
```
```{r 16-dynamic-labeling-to-variable}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_text(aes(label = drv))
```

When adding more than one `geom()`-function, multiple geometries are added to the plot.
However, because `geom_point()` has no support for passing a label, we can only use this mapping in `geom_text()`.

```{r include = FALSE}
# points and labels overplotting
# only geom_text() supports label arg
```
```{r 16-points-and-labels-overplotting}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  geom_text(aes(label = drv))
```

Since this looks just slightly odd, let's try to make it more apparent, what is happening.

```{r include = FALSE}
# aesthetic changes to distinguish better
```
```{r 16-aesthetic-changes-to-distinguish-better}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point(color = "blue") +
  geom_text(aes(label = drv), size = 10)
```
