```{r 17-remove-all, include = FALSE}
rm(list = ls())
```


## Themes

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Themes

```
```{r 17-setup-17-theme-r}
library(tidyverse)
```
</details>

<br>

In this section we are looking at the use of visual themes to easily change the look and feel of a plot.
We start with the introduction of the default theme -- `theme_grey()` function.

```{r include = FALSE}
# Base plot with default theme
```
```{r 17-base-plot-with-default-theme}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  theme_grey()
```

Change the default `theme_grey()` to a more traditional black-and-white theme:

```{r include = FALSE}
# bw-theme
```
```{r 17-bw-theme}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point() +
  theme_bw()
```

Also in this scheme the color aesthetic works as it normally does.
The black-and-whiteness only relates to the background.

```{r include = FALSE}
# Theme does not affect color palette of geoms
```
```{r 17-theme-does-not-affect-color-palette-of-geoms}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point(aes(color = class)) +
  theme_bw()
```

Calling the function `theme()` after a `theme_...()` call let's you tweak certain aspects of the theme.

Some plots work better with the legend at the bottom.

```{r include = FALSE}
# Occasionally useful: legend at the bottom
```
```{r 17-occasionally-useful-legend-at-the-bottom}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point(color = class) +
  theme_bw() +
  theme(legend.position = "bottom")
```

Mind that `theme_...()` functions overwrite all previous settings of `theme()`:

```{r include = FALSE}
# theme_* overwrites theme() calls: mind the order!
```
```{r 17-theme-overwrites-theme-calls-mind-the-order}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point(color = class) +
  theme(legend.position = "bottom") +
  theme_bw()
```

The first argument of each `theme_...()` function is `base_size`, which refers to the font size of all elements in the plot.

```{r include = FALSE}
# first param of theme_()-functions is base_size
```
```{r 17-first-param-of-theme-functions-is-base-size}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point() +
  theme_bw(16)
```

If we were asked to suggest themes, we'd go for 

- `ggplot2::theme_minimal()` 
- `hrbrthemes::theme_ipsum()`
- `ggpubr::theme_pubr()`

Here is how `ggpubr::theme_pubr()` looks like.

```{r include = FALSE}
# predefined custom themes available, e.g. in ggpubr package
```
```{r 17-predefined-custom-themes-available-e-g-in-ggpubr-package}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point() +
  ggpubr::theme_pubr()
```

Also from here onward we will use `theme_pubr()` as the default theme for plots.
This can be done by setting 
