## Scales

<details><summary>*Click here to show setup code.*</summary>
```{r 19-setup-19-scales-r}
library(tidyverse)
library(nycflights13)
library(ggpubr)

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```
</details>

<br>

In this section we want to spend some time getting to know how to customize the labels and scales of plots using {ggplot2}.
We start with a pretty basic plot using the `mpg`-tibble which comes with the {tidyverse}.

```{r 19-base-plot}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point() +
  theme_pubr()
```

### `labs()`

With `labs()` you can label all sorts of aesthetics (axes, color mapping, ...). 
Additionally you can set the title/subtitle and also add a caption and a tag.

```{r 19-add-descriptive-labels-the-labs-function}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point() +
  theme_pubr() +
  labs(
    x = "Displacement",
    y = "Highway mileage\n[miles per gallon]",
    color = "Car class",
    title = "Highway mileages depending on displacement",
    subtitle = "By car class"
  )
```

### Axes

There is a plethora of `scale_...()` functions available in {ggplot2}, which influence the axes.
For example there is a function to change the scale of an axis to a logarithmic scale.

```{r 19-change-scale-to-log-scale}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point() +
  theme_pubr() +
  scale_x_log10()
```

Be careful: you can set the name of an axis in both the `labs()` function and the `scale_...()` functions.
If you do both, only the name set in the latter will prevail.

```{r 19-name-in-scale-overwrites-labs-without-warning}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point() +
  labs(
    x = "Displacement"
  ) +
  theme_pubr() +
  scale_x_log10(name = "xxx")
```

---

For more control over discrete and continuous axis labels, limits and breaks, the `scale_<axis name>_<variable type>` functions exist, e.g. `scale_x_continuous()`.

These enable custom axis breaks and labels if the ones autogenerated from the data are not sufficient.

```{r 19-custom-axis-breaks}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  theme_pubr() +
  scale_x_continuous(limits = c(2, 6), breaks = c(2, 4, 6))
```

Values not falling into the custom limits will be silently droppend including a warning message.

### Color scale

Another type `scale_...()`-type function relates to the color-aesthetic.
These functions affect the palette that is used for the color mapping.
By default, `scale_color_hue()` will be used for categorical variables.

```{r 19-default}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
) +
  geom_point(aes(color = class)) +
  theme_pubr() +
  scale_color_hue()
```

To change the color palette, pass a palette-function of your liking in the form of 

- `scale_color_<name>` 
- `scale_fill_<name>`

Whether to use `fill` or `color` depends on what keyword has been used for applying the color.
Points are colored by using the "color" keyword.
So to change the palette for point coloring, one needs to use `scale_color_<name>`.

A popular color palette is the [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html#gallery) color palette. 
To specify that we are dealing with categorical values, we add a `_d` at the end which stands for "discrete".

```{r 19-other-options-viridis}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point(aes(color = class)) +
  theme_pubr() +
  scale_color_viridis_d()
```

---

To take full control of the colors `scale_color_manual()` should be used.
Here, color values (either as a string or in hex format) can be bound to a specific factor level.

This is useful if certain levels come with implicit meanings of their color choice.
Another helpful scenario is when there are more levels in the data than colors supported by the palette (most palettes support between 9-12 levels).

```{r 19-full-control-manual-scale}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = class)
) +
  geom_point() +
  theme_pubr() +
  scale_color_manual(values = c(
    "2seater" = "#000000",
    "compact" = "#3355FF",
    "midsize" = "#006400",
    "minivan" = "#FF5522",
    "pickup" = "#66FFFF",
    "subcompact" = "#FF0000",
    "suv" = "#FF55FF"
  ))
```

---

Review the [{ggthemr} package](https://github.com/cttobin/ggthemr) for tools that help with establishing a "corporate design" for documents.

```{r 19-for-corporate-design-ggthemr-https-github-com-cttobin-ggthemr, eval=FALSE}
install.packages("remotes")
remotes::install_packages("cttobin/ggthemr")
```

**Exercises**

https://krlmlr.github.io/vistransrep/2019-11-zhr/scales.html
