```{r 206-remove-all, include = FALSE}
rm(list = ls())
```


## Create new columns

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Create new columns

```
```{r 206-setup-206-mutate-r}
library(tidyverse)
library(nycflights13)

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```
</details>

With `dplyr::mutate()` you can add new columns to a table, e.g. making use of the already existing variables.
This is another building block added to the toolset.

How much faster than the scheduled time did the pilots manage to fly:

```{r include = FALSE}
# Compute new variables
```
```{r 206-compute-new-variables}
flights %>%
  mutate(recovery = dep_delay - arr_delay) %>%
  select(dep_delay, arr_delay, recovery)
```

Conceptually, the expression that defines the new variable is evaluated for each row.

The following constructs are often applied inside `mutate()`:

- Arithmetic: `+`, `-`, `*`, `/`, `^`, `%%`, `%/%`

    ```r
    dep_delay - arr_delay
    dep_time %/% 100
    dep_time %% 100
    dep_delay - mean(dep_delay)   # Deviation from mean
    ```

- Real functions, see `?base::Math` and `?dplyr::lead`:
    - Rounding: `floor()`, `ceiling()`, `round()`
    - Sign: `abs()`, `sign()`
    - Transform: `sqrt()`, `log()`, `log2()`, `exp()`
    - Trigonometric: `sin()` etc.
    
- Recoding: `if_else()`, `case_when()`, `recode()`

- All filtering functions for a new `logical` column

- `str_replace()` for string columns

- Functions that process values from other rows:
    - Cumulative: `cumsum()` etc.
    - Lead and lag: `lead()`, `lag()`
    - Ranking: `row_number()`, `min_rank()`, `ntile()`

Work with the newly created variables just like with the original ones:

```{r include = FALSE}
# Work with new variables
```
```{r 206-work-with-new-variables}
flights %>%
  mutate(recovery = dep_delay - arr_delay) %>%
  select(dep_delay, arr_delay, recovery) %>%
  arrange(recovery)
```

A `mutate()` never changes a dataset.
To make a computation persistent, store the entire result as a new dataset variable.

```{r include = FALSE}
# Assign to persist:
```
```{r 206-assign-to-persist}
try(
  flights %>%
  select(recovery)
)

recovery_data <-
  flights %>%
  mutate(recovery = dep_delay - arr_delay) %>%
  select(dep_delay, arr_delay, recovery) %>%
  arrange(recovery)

recovery_data
```

Let's look at a single airplane:

```{r include = FALSE}
# flights filter select view
```
```{r 206-flights-filter-select-view}
flights %>%
  filter(tailnum == "N14228") %>%
  select(year, month, day, dep_time, arr_time) %>%
  view()
```

Adding the departure time of the *next* flight to the current row, respectively, using `mutate()` with `lead()`:

```{r include = FALSE}
# Lead values look across rows:
```
```{r 206-lead-values-look-across-rows}
flights %>%
  filter(tailnum == "N14228") %>%
  select(year, month, day, dep_time, arr_time) %>%
  mutate(lead_dep_time = lead(dep_time)) %>%
  view()
```

The opposite effect to `lead()` can be realized using `lag()`:

```{r include = FALSE}
# Lag: opposite of lead:
```
```{r 206-lag-opposite-of-lead}
flights %>%
  filter(tailnum == "N14228") %>%
  select(year, month, day, dep_time, arr_time) %>%
  mutate(lag_arr_time = lag(arr_time)) %>%
  view()
```

There is even a use-case for this in our little example.
How long does it take for the airplane to return to NYC with each flight out?

```{r include = FALSE}
# Compute absence time from NYC airports:
```
```{r 206-compute-absence-time-from-nyc-airports}
flights %>%
  filter(tailnum == "N14228") %>%
  select(year, month, day, time_hour) %>%
  mutate(lag_time_hour = lag(time_hour)) %>%
  mutate(ground_time = time_hour - lag_time_hour) %>%
  view()
```

A frequently used workflow is creating a helper variable at some point in the pipeline and then dropping it later on:

```{r include = FALSE}
# Create helper, then deselect with `-`:
```
```{r 206-create-helper-then-deselect-with}
flights %>%
  filter(tailnum == "N14228") %>%
  select(year, month, day, dep_time, arr_time) %>%
  mutate(lag_arr_time = lag(arr_time)) %>%
  mutate(ground_time = dep_time - lag_arr_time) %>%
  select(-lag_arr_time)
```

Let's work some more with the flight data of our special plane.
The total air time of a plane up to and including a given flight can be calculated with `cumsum()`:

```{r include = FALSE}
# Cumulative air time:
```
```{r 206-cumulative-air-time}
flights %>%
  filter(tailnum == "N14228") %>%
  mutate(cum_air_time = cumsum(air_time)) %>%
  select(air_time, cum_air_time) %>%
  view()
```

Creating a "flag" variable with `mutate()` allows selecting if a flight was on time or not:

```{r include = FALSE}
# Verbose indicators:
```
```{r 206-verbose-indicators}
flights %>%
  mutate(delayed = if_else(arr_delay > 0, "delayed", "on time")) %>%
  select(arr_delay, delayed)
```

Shorter, but less verbose:

```{r include = FALSE}
# Logical flags:
```
```{r 206-logical-flags}
flights %>%
  mutate(delayed = (arr_delay > 0)) %>%
  select(arr_delay, delayed)
```

A flag can be passed on to `filter()` directly:

```{r include = FALSE}
# Filter by flag:
```
```{r 206-filter-by-flag}
flights %>%
  mutate(delayed = arr_delay > 0) %>%
  select(arr_delay, delayed) %>%
  filter(delayed)
```

Use negation for inverse filtering, and store in a dataset variable for reuse:

```{r include = FALSE}
# Negation, persistence:
```
```{r 206-negation-persistence}
on_time_flights <-
  flights %>%
  mutate(delayed = (arr_delay > 0)) %>%
  select(arr_delay, delayed) %>%
  filter(!delayed)

on_time_flights
```

### Exercises {-}

https://krlmlr.github.io/vistransrep/2019-11-zhr/mutate.html
