```{r 209-remove-all, include = FALSE}
rm(list = ls())
```


## Joins

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Joins

```
```{r 209-setup-209-join-r}
library(tidyverse)
library(nycflights13)

library(here)

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
set.seed(20191129)
```
</details>


Joins combine two tables horizontally by combining matching rows.
Two rows match if they have the same value in columns common to both tables.
The resulting table contains all columns from both tables.

The `airlines` data contains details (really just the name) of airlines, by carrier code.

```{r include = FALSE}
# `airlines` contain detailed information for each airline
# (in this case only the name)
```
```{r 209-airlines-contain-detailed-information-for-each-airline}
airlines
```

Only the carrier code is available in the `flights` dataset.
The airline name can be brought in with the `left_join()` function, e.g., to use in a pretty label.

```{r include = FALSE}
# `flights` can be augmented with this detailed information
```
```{r 209-flights-can-be-augmented-with-this-detailed-information}
flights %>%
  count(carrier) %>%
  left_join(airlines)
```

For code stability, it is useful to always specify the column names to join by.

```{r include = FALSE}
# for "production code", always indicate the columns to join by
```
```{r 209-for-production-code-always-indicate-the-columns-to-join-by}
flights %>%
  left_join(airlines, by = "carrier") %>%
  select(dep_time, carrier, name)
```


** Your turn**: Bring in data from the `planes` table.
What happens if you omit the `by` argument?

```{r include = FALSE}
# Your turn: join the `planes` table
```
```{r 209-your-turn-join-the-planes-table}
planes
```


```{r include = FALSE}
# Does it work out of the box?
```
```{r 209-does-it-work-out-of-the-box}
try(
  ... %>%
    ..._join(...)
)
```


```{r include = FALSE}
# How to fix, and also turn off the message?
```
```{r 209-how-to-fix-and-also-turn-off-the-message}
try(
  ... %>%
    ..._join(..., ... = "...")
)
```

The `airports` table does not have any columns common to `flights`.

```{r include = FALSE}
# `airports` contain detailed information for each airport
```
```{r 209-airports-contain-detailed-information-for-each-airport}
airports

try(
  flights %>%
    left_join(airports)
)
```

A `by` argument with a specific syntax is required here.

```{r include = FALSE}
# in this case, explicit "by" columns are required
```
```{r 209-in-this-case-explicit-by-columns-are-required}
flights %>%
  left_join(airports, by = c("dest" = "faa")) %>%
  select(origin, dest, name)
```

Mismatches introduce cells with missing values.

```{r include = FALSE}
# we have mismatches!
```
```{r 209-we-have-mismatches}
flights %>%
  left_join(airports, by = c("dest" = "faa")) %>%
  select(origin, dest, name) %>%
  filter(is.na(name))
```

The `airports` table does not contain a row for the BQN airport.

```{r include = FALSE}
# what about the NA name?
```
```{r 209-what-about-the-na-name}
airports %>%
  filter(faa == "BQN")
```

The `inner_join()` discards rows with mismatches.
Beware: It is tempting to always use the inner join, but often a mismatch indicates problems that have occurred earlier in your analysis.

```{r include = FALSE}
# left vs. inner vs. outer joins control what happens in case of mismatches
```
```{r 209-left-vs-inner-vs-outer-joins-control-what-happens-in-case-of-mismatches}
flights %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  select(origin, dest, name)
```


**Your turn**: Why are some rows with the `planes` table not matched?

```{r include = FALSE}
# Your turn: understand mismatches when joining the `planes` data
```
```{r 209-your-turn-understand-mismatches-when-joining-the-planes-data}
try(
  ... %>%
    left_join(..., ... = "...") %>%
    filter(is.na(...))
)
```


```{r include = FALSE}
# Verify that "AA" has more mismatches than other airlines
```
```{r 209-verify-that-aa-has-more-mismatches-than-other-airlines}
try(
  ... %>%
    left_join(..., ... = "...") %>%
    mutate(mismatch = is.na(...), is_aa = (carrier == "AA")) %>%
    count(.....)
)
```


```{r include = FALSE}
# How to keep only the flights where we have airplane details?
```
```{r 209-how-to-keep-only-the-flights-where-we-have-airplane-details}
try(
  ... %>%
    ..._join(..., ... = "...")
)
```

The following example uses `rpart::rpart()` to classify missingness by the other columns.
Turns out there is a pattern, which needs to be investigated further in the original data.

```{r include = FALSE}
# Case study: classification of mismatches
```
```{r 209-case-study-classification-of-mismatches}
flights %>%
  left_join(planes %>% select(tailnum, manufacturer), by = "tailnum") %>%
  mutate(mismatch = is.na(manufacturer)) %>%
  select(-tailnum, -manufacturer) %>%
  rpart::rpart(mismatch ~ ., .)
```

For complex joins, it is worthwhile preparing the tables you intend to join beforehand.

```{r include = FALSE}
# before joining, prepare the RHS table(s)
```
```{r 209-before-joining-prepare-the-rhs-table-s}
origin_airports <-
  airports %>%
  select(origin = faa, origin_name = name)

origin_airports

dest_airports <-
  airports %>%
  select(dest = faa, dest_name = name)

dest_airports
```


```{r include = FALSE}
# much easier with the prepared tables
```
```{r 209-much-easier-with-the-prepared-tables}
flights %>%
  left_join(origin_airports) %>%
  left_join(dest_airports) %>%
  select(origin, origin_name, dest, dest_name)
```

Still, it is recommended to explicitly specify the `by` column.

```{r include = FALSE}
# "by" still recommended
```
```{r 209-by-still-recommended}
flights %>%
  left_join(origin_airports, by = "origin") %>%
  left_join(dest_airports, by = "dest") %>%
  select(origin, origin_name, dest, dest_name)
```


**Your turn**: How do we bring in only the `engines` and `seats` columns from the `planes` table?

```{r include = FALSE}
# Your turn: we only need `engines` and `seats` from the `planes` table
```
```{r 209-your-turn-we-only-need-engines-and-seats-from-the-planes-table}
try({
  planes_join <-
    ... %>%
    ...(...)

  ... %>%
    left_join(..., ... = "...")
})
```

### Keys

For the join to work as expected, it is important that the columns in `by` uniquely define observations, either in the left-hand-side or in the right-hand-side table, or in both.
If this is given, the column or columns are a *primary key* to the table.
The `count()` function helps asserting this.

```{r include = FALSE}
# either LHS or RHS should have a key as part of "by"
```
```{r 209-either-lhs-or-rhs-should-have-a-key-as-part-of-by}
airports %>%
  count(faa)

airports %>%
  count(faa) %>%
  count(n)
```

**Your turn**: Double-check that `tailnum` is indeed a key to `planes`.

```{r include = FALSE}
# Your turn: double-check that `tailnum` is indeed a key in `planes`
```
```{r 209-your-turn-double-check-that-tailnum-is-indeed-a-key-in-planes}
try(
  ... %>%
    count(...) %>%
    ...(...)
)
```

In the example below we derive a dataset from `airports` that intentionally has duplicate observations.
The `faa` column is no longer a key to the table.

```{r include = FALSE}
# create artificial dataset where airport is not a key
```
```{r 209-create-artificial-dataset-where-airport-is-not-a-key}
dup_airports <-
  bind_rows(airports, sample_n(airports, 100))

dup_airports %>%
  count(faa)

dup_airports %>%
  count(faa) %>%
  count(n)

dup_airports %>%
  count(faa) %>%
  add_count(n) %>%
  filter(n > 1)
```

Because we have duplicate observations with the same value in the `faa` column (and of course we have many more flights), the join will produce duplicate flights.
This is almost always undesirable!

```{r include = FALSE}
# what happens if not key?
```
```{r 209-what-happens-if-not-key}
flights %>%
  left_join(dup_airports, by = c("dest" = "faa")) %>%
  select(origin, dest, name)

flights
```

**Your turn**: Find a key to the `weather` table.

```{r include = FALSE}
# Your turn: Is this combination a key? Why/why not?
```
```{r 209-your-turn-is-this-combination-a-key-why-why-not}
weather

weather %>%
  count(origin, year, month, day, hour) %>%
  count(n)
```


```{r include = FALSE}
# How to find the offenders?
```
```{r 209-how-to-find-the-offenders}
try(
  weather %>%
    count(...) %>%
    add_count(...) %>%
    filter(...)
)
```


```{r include = FALSE}
# Explain!
```
```{r 209-explain}

```

The [{dm} package](https://krlmlr.github.io/dm/) is a modern take on handling multiple tables in your R session.
A `dm` object combines the data and metadata for tables, and the relationships between these tables.
Joins no longer require specification of the `by` columns.
The join logic becomes part of the metadata and not the code.

```{r include = FALSE}
# A novel approach: https://krlmlr.github.io/dm/
```
```{r 209-a-novel-approach-https-krlmlr-github-io-dm}

```


```{r include = FALSE}
##install.packages("devtools")
##devtools::install_github("krlmlr/dm")
```
```{r 209-install-packages-devtools}
NA
```
