```{r 23-remove-all, include = FALSE}
rm(list = ls())
```


## Sort rows: `dplyr::arrange()`

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Sort rows: `dplyr::arrange()`

```
```{r 23-setup-23-arrange-r}
library(tidyverse)
library(nycflights13)

library(conflicted)
conflict_prefer("filter", "dplyr")
```
</details>


The function `dplyr::arrange()` sorts the rows of the dataset according to the values of the variable(s) you are providing.

```{r include = FALSE}
# simple arrange
```
```{r 23-simple-arrange}
flights %>%
  arrange(dep_time)
```

When providing multiple variables as arguments for `...` (the ellipsis), the dataset is first sorted accorcing to the values of the first variable.
Wherever these values occur more than once, another sorting takes place within those groups, according to the second variable you provided.
The same rule applies for every further variable you add to `arrange()`.

```{r include = FALSE}
# lexicographically arranging two columns:
# first column, resolve ties with the second column
```
```{r 23-lexicographically-arranging-two-columns}
flights %>%
  arrange(dep_time, dep_delay)
```

You can combine `filter()` and `arrange()`.

```{r include = FALSE}
# filtering and arranging can be combined
```
```{r 23-filtering-and-arranging-can-be-combined}
flights %>%
  filter(dep_time < 600) %>%
  filter(month >= 10) %>%
  arrange(dep_time, dep_delay) %>%
  view()
```

You can use `arrange()` with arbitrary expressions.

```{r include = FALSE}
# moving NA values to end
```
```{r 23-moving-na-values-to-end}
flights %>%
  filter(month == 4) %>%
  filter(day == 1) %>%
  arrange(is.na(dep_time)) %>%
  view()
```

The reason for the result you just saw in the view of the filtered dataset is, that the binary result of the expression (`TRUE`, `FALSE`) is sorted `FALSE` first (lexicographically).

Let's give it a twist:

```{r include = FALSE}
# moving NA values to start
```
```{r 23-moving-na-values-to-start}
flights %>%
  filter(month == 4) %>%
  filter(day == 1) %>%
  arrange(!is.na(dep_time)) %>%
  view()
```

Sorting the dataset according to which flights arrived earliest on April 1, 2013:

```{r include = FALSE}
# two filters then arrange according arr_time
```
```{r 23-two-filters-then-arrange-according-arr-time}
flights %>%
  filter(month == 4) %>%
  filter(day == 1) %>%
  arrange(arr_time) %>%
  view()
```

Invert the sorting by either...

```{r include = FALSE}
# Descending
```
```{r 23-descending}
flights %>%
  filter(month == 4) %>%
  filter(day == 1) %>%
  arrange(-arr_time) %>%
  view()
```

... or:

```{r include = FALSE}
# Descending 2
```
```{r 23-descending-2}
flights %>%
  filter(month == 4) %>%
  filter(day == 1) %>%
  arrange(desc(arr_time)) %>%
  view()
```

You can mix sorting in an ascending and a descending manner:

```{r include = FALSE}
# Ascending and descending
```
```{r 23-ascending-and-descending}
flights %>%
  filter(month == 4) %>%
  filter(day == 1) %>%
  arrange(dep_time, desc(arr_time)) %>%
  view()
```
