```{r 24-remove-all, include = FALSE}
rm(list = ls())
```


## The pipe

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### The pipe

```
```{r 24-setup-24-pipe-r}
library(tidyverse)
library(nycflights13)

library(conflicted)
conflict_prefer("filter", "dplyr")
view <- I
```
</details>

We already heavily used it today, but what exactly are the characteristics of `%>%`, better known as "the pipe"?

```{r include = FALSE}
# early flights
```
```{r 24-early-flights}
early_flights <-
  flights %>%
  filter(dep_time < 600)
```

The above is just another way of writing:

```{r include = FALSE}
# early flights with pipe
```
```{r 24-early-flights-with-pipe}
early_flights <- filter(flights, dep_time < 600)
```

You can consult the manual:

```{r include = FALSE}
# pipe help
```
```{r 24-pipe-help}
?"%>%"
```

What are the benefits of using the pipe?
E.g. you can swap the order of the rows with `Alt + Cursor up/down` easily, if you want to exchange the building blocks for your analysis:

```{r include = FALSE}
# Alt + Cursor up/down
```
```{r 24-alt-cursor-up-down}
flights %>%
  filter(dep_time < 600) %>%
  filter(month >= 10) %>%
  arrange(dep_time, dep_delay) %>%
  view()
```

When working on a code chunk consisting of subsequent transformations connected by pipes, it can be useful to end the pipeline with either `I` or `view()`.

```{r include = FALSE}
# Add/remove I or view
```
```{r 24-add-remove-i-or-view}
flights %>%
  filter(dep_time < 600) %>%
  filter(month >= 10) %>% I
```

But once the chunk does what you expect it to do, do not forget to remove the `I` or `view()` call.

```{r include = FALSE}
# Remember to remove it!
```
```{r 24-remember-to-remove-it}
  # arrange(dep_time, dep_delay) %>%
  # view()
```
