```{r 24-remove-all, include = FALSE}
rm(list = ls())
```


## The pipe

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### The pipe

```
```{r 24-setup-24-pipe-r}
library(tidyverse)
library(nycflights13)

library(conflicted)
conflict_prefer("filter", "dplyr")
```
</details>

We already heavily used it today, but what exactly are the characteristics of `%>%`, better known as "the pipe"?

```{r include = FALSE}
# early flights
```
```{r 24-early-flights}
early_flights <-
  flights %>%
  filter(dep_time < 600)
```

The above is just another way of writing:

```{r include = FALSE}
# early flights with pipe
```
```{r 24-early-flights-with-pipe}
early_flights <- filter(flights, dep_time < 600)
```

The manual describes this operator in detail:

```{r include = FALSE}
# pipe help
```
```{r 24-pipe-help, eval=FALSE}
?"%>%"
```

With the pipe, code can be read in a natural way, from left to right.
The following snippet extracts

1. all early flights
2. from October till December,
3. ordered by departure time and then departure delay
4. and displays it.

Note how the reading corresponds to the code.

```{r include = FALSE}
# Complex pipe
```
```{r 24-complex-pipe}
flights %>%
  filter(dep_time < 600) %>%
  filter(month >= 10) %>%
  arrange(dep_time, dep_delay) %>%
  view()
```

This is possible, because all transformation verbs (`filter()`, `arrange()`, `view()`) accept the main input (a tibble) as the first argument and also return a tibble.

The following three codes are equivalent, but are more difficult to write, to read and to maintain.

Naming is hard.
Trying to give each intermediate result a name is exhausting.
Introducing an additional step in this sequence of operations is prone to errors.

```{r include = FALSE}
# Intermediate names
```
```{r 24-intermediate-names}
early_flights <- filter(flights, dep_time < 600)
early_flights_oct_dec <- filter(early_flights, month >= 10)
early_flights_oct_dec_sorted <- arrange(early_flights_oct_dec, dep_time, dep_delay)
view(early_flights_oct_dec_sorted)
```

We can keep using the same variable, e.g. `x`, to avoid naming.
This adds noise compared to the pipe.

```{r include = FALSE}
# Catch-all variable
```
```{r 24-catch-all-variable}
x <- flights
x <- filter(x, dep_time < 600)
x <- filter(x, month >= 10)
x <- arrange(x, dep_time, dep_delay)
view(x)
```

We can avoid intermediate variables.
This disconnects the verbs from their arguments and is very difficult to read.

```{r include = FALSE}
# Layers
```
```{r 24-layers}
view(
  arrange(
    filter(
      filter(
        flights,
        dep_time < 600
      ),
      month >= 10
    ),
    dep_time, dep_delay
  )
)
```


### Further advantages

When working on a code chunk consisting of subsequent transformations connected by pipes, it can be useful to end the pipeline with either `I` or `view()`.

```{r include = FALSE}
# Add/remove I or view
```
```{r 24-add-remove-i-or-view}
flights %>%
  filter(dep_time < 600) %>%
  filter(month >= 10) %>% I
##arrange(dep_time, dep_delay) %>%
##view()
```

Once the chunk does what you expect it to do, do not forget to remove the `I` or `view()` call.

```{r include = FALSE}
# Remember to remove it!
```
```{r 24-remember-to-remove-it}
try(
  arrange(dep_time, dep_delay) %>%
  view()
)
```

To rearrange rows, you can use the shortcut `Alt + Cursor up/down`.
In a piped expression, no further editing is necessary!
