```{r 25-remove-all, include = FALSE}
rm(list = ls())
```


## Pick columns: `dplyr::select()`

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Pick columns: `dplyr::select()`

```
```{r 25-setup-25-select-r}
library(tidyverse)
library(nycflights13)

library(conflicted)
conflict_prefer("filter", "dplyr")
```
</details>

With `dplyr::select()` you can (de-)select and/or rename columns of your dataset.
The basic operation is like in the following examples:

```{r include = FALSE}
# select columns
```
```{r 25-select-columns}
flights %>%
  select(year, month, day)
```


```{r include = FALSE}
# select a column away
```
```{r 25-select-a-column-away}
flights %>%
  select(-year)
```

Renaming works by addressing an existing column on the right hand side of an equality sign and providing the new name of the column on its left hand side.



```{r include = FALSE}
# select supports renaming
```
```{r 25-select-supports-renaming}
flights %>%
  select(
    year, month, day,
    departure_delay = dep_delay,
    arrival_delay = arr_delay
  )
```


With backticks, it is possible, but not advised, to use arbitrary characters (including spaces) in column names:

```{r include = FALSE}
# select + rename: create spaces (don't do this!)
```
```{r 25-select-rename-create-spaces-don-t-do-this}
flights_with_spaces <-
  flights %>%
  select(
    year, month, day,
    `Departure delay` = dep_delay,
    `Arrival delay` = arr_delay
  ) %>%
  filter(
    `Arrival delay` < 0
  )
```

Address them in the same way, if the dataset already has such variables:

```{r include = FALSE}
# select + rename columns with spaces
```
```{r 25-select-rename-columns-with-spaces}
flights_with_spaces %>%
  select(
    year, month, day,
    dep_delay = `Departure delay`,
    arr_delay = `Arrival delay`
  )
```

The {janitor} package helps fixing issues with colum names automatically.

Select helpers allow selecting multiple related columns conveniently.

```{r include = FALSE}
# select helpers
```
```{r 25-select-helpers}
flights %>%
  select(origin, dest, ends_with("_time"))
```
