```{r 28-remove-all, include = FALSE}
rm(list = ls())
```


## Summary-plots

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Summary-plots

```
```{r 28-setup-28-multiplot-r}
library(tidyverse)
library(nycflights13)

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```
</details>

Potentially surprisingly, `mutate()` can also work with the results of a `ggplot()` call.
Let's approach this step by step.
Here is a basic barplot of `flights$carrier`:

```{r include = FALSE}
# Bar plot: flights per carrier
```
```{r 28-bar-plot-flights-per-carrier}
flights %>%
  ggplot(aes(x = carrier)) +
  geom_bar()
```

Same with one facet per month:

```{r include = FALSE}
# Bar plot: flights per carrier, one facet per month
```
```{r 28-bar-plot-flights-per-carrier-one-facet-per-month}
flights %>%
  ggplot(aes(x = carrier)) +
  geom_bar() +
  facet_wrap(~month)
```

We can extract a function that takes any data and produces a barplot of the variable `carrier`:

```{r include = FALSE}
# Extract a function
```
```{r 28-extract-a-function}
plot_fun <- function(data) {
  data %>%
    ggplot(aes(x = carrier)) +
    geom_bar()
}

plot_fun(flights)
```

The result of `ggplot()` is first and foremost an object.
Only when R tries to display it on the console a method is triggered, which causes it to show the graph in the "Viewer".
Therefore, we can use the `group_by` -- `summarize()` -- `ungroup()` pattern to produce one plot per group and store it in a new column:

```{r include = FALSE}
# Call this function for each subgroup
```
```{r 28-call-this-function-for-each-subgroup}
plot_df <-
  flights %>%
  group_by(month) %>%
  summarize(
    plot = list(plot_fun(tibble(carrier)))
  ) %>%
  ungroup()

plot_df
```

When using `dplyr::pull()` (this function "extracts" a variable from a `data.frame` and returns it as a normal vector), each of the plots will be subsequently displayed in your "Viewer".

```{r include = FALSE}
# Show the plots
```
```{r 28-show-the-plots}
plot_df %>%
  pull()
```

Use the left arrow to click through the different plots.
