```{r 32-remove-all, include = FALSE}
rm(list = ls())
```


## Import many files

<details><summary>*Click here to show setup code.*</summary>
```{r include = FALSE}
### Import many files

```
```{r 32-setup-32-import-many-r}
library(tidyverse)
library(nycflights13)

library(here)

library(conflicted)
conflict_prefer("filter", "dplyr")
```
</details>

Occasionally, a dataset is split across many files with a very similar format.
The `data/` directory contains several Excel files with the `.xlsx` extension with tables of nearly identical format.

```{r include = FALSE}
# List all files of interest
```
```{r 32-list-all-files-of-interest}
files <- dir(path = here("data"), pattern = "[.]xlsx$", full.names = TRUE)
files
```

An easy way to import all files at once is the `rio::import_list()` function from the {rio} package.

```{r include = FALSE}
# Import all files at once with rio
```
```{r 32-import-all-files-at-once-with-rio}
files %>%
  rio::import_list(setclass = class(tibble()), rbind = TRUE)
```

If some files need manipulation before the data can be bound together, {rio} also offers a way to import them as a "named list".

```{r include = FALSE}
# Import into named list first
```
```{r 32-import-into-named-list-first}
list_of_tables <- rio::import_list(files, setclass = class(tibble()))
list_of_tables
```

The data can be accessed individually for each input file.

```{r include = FALSE}
# Access individual tables
```
```{r 32-access-individual-tables}
list_of_tables$example6b
```

If a tweak is necessary, the data can be overwritten as needed.

```{r include = FALSE}
# Maybe tweak individual files?
```
```{r 32-maybe-tweak-individual-files}
try(
  list_of_tables$example6b <-
    list_of_tables$example6b %>%
    mutate(...) %>%
    select(...)
)
```

The `bind_rows()` function combines these components into a single dataset again.

```{r include = FALSE}
# Combine into a single data frame
```
```{r 32-combine-into-a-single-data-frame}
all_tables <- bind_rows(list_of_tables, .id = "path")
all_tables
```

When done, use `filter()` to access a single dataset.

```{r include = FALSE}
# Analyze one file
```
```{r 32-analyze-one-file}
all_tables %>%
  filter(path == "example6b") %>%
  summarize(mean(col1), first(col2))
```

For performing an analysis across the entire dataset, per input file, use `group_by()`:

```{r include = FALSE}
# Analyze multiple files: use group_by()
```
```{r 32-analyze-multiple-files-use-group-by}
all_tables %>%
  group_by(path) %>%
  summarize(mean(col1), first(col2)) %>%
  ungroup()
```

Finally, `map_dfr()` offers a way to import files with more control.
The details are out of scope here.

```{r include = FALSE}
# Alternative: Import manually
```
```{r 32-alternative-import-manually}
files %>%
  map_dfr(~ readxl::read_excel(.))
```

### Exercises {-}

https://krlmlr.github.io/vistransrep/2019-11-zhr/import.html, task 3 and 4
