---
title: "Position adjustments"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Gears

The information is available in the `trans` attribute.

```{r}
ggplot(mpg, aes(trans)) +
  geom_bar()
```

Visual inspection reveals that it's four gears for automatic, and five gears for manual transmission.


## Discriminate by car class

```{r}
ggplot(mpg, aes(class, fill = trans)) +
  geom_bar()
```

SUVs have automatic mostly, 2-seaters favor six-gear manual, compact and subcompact mostly have five-gear manual transmission. This is much easier to see with `position = "dodge"`, the different width of the bars indicates the number different subgroups for each car class:

```{r}
ggplot(mpg, aes(class, fill = trans)) +
  geom_bar(position = "dodge")
```

## Boxplot

I'm using liters per 100 km as measure for fuel economy here.

```{r}
ggplot(mpg, aes(drv, 235 / hwy)) +
  geom_boxplot()
```

Forward drivetrains seem much more economical. Does the number of cylinders play a role? I'll try the "fill" aesthetic:

```{r}
ggplot(mpg, aes(drv, 235 / hwy, fill = cyl)) +
  geom_boxplot()
```

No dice. Do I also need `group =` ?

```{r}
ggplot(mpg, aes(drv, 235 / hwy, fill = cyl, group = cyl)) +
  geom_boxplot()
```

The legend reveals that `cyl` is a continuous variable. I'll use its categorical equivalent, because the range is very limited.

```{r}
ggplot(mpg, aes(drv, 235 / hwy, fill = factor(cyl))) +
  geom_boxplot(position = "identity")
```

## Summary

The default position adjustments were fine for the assignments in this exercise.
