---
title: "Statistical transformations"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Bar plot

I tried `geom_histogram()` and `geom_col()`, neither worked. The histogram is for continuous data only, for `geom_col()` I'd need to supply actual counts which I don't have. The `geom_bar()` function computes the counts for me by applying the `"count"` statistical transformation to my data before plotting.

```{r}
ggplot(mpg) +
  geom_bar(aes(x = drv))
```


## ..count..

I think the results look pretty much the same if I add `y = ..count..` to the `aes()` call.
Perhaps this is related to `count` being a "computed variable" (according to `?geom_bar`)?


```{r}
ggplot(mpg) +
  geom_bar(aes(x = drv, y = ..count..))
```

## ..prop..



## Fuel economy

The histogram call was pretty straightforward, very similar to the bar chart above.

```{r}
ggplot(mpg) +
  geom_histogram(aes(hwy))
```

I did obey the warning and specified a width for the bins:

```{r}
ggplot(mpg) +
  geom_histogram(aes(hwy), binwidth = 2)
```

To overlay the frequency polygon, I use a second layer. I can make the aesthetic mapping global because both layers use the same.

```{r}
ggplot(mpg, aes(hwy)) +
  geom_histogram(binwidth = 2) +
  geom_freqpoly()
```

Can I specify the bin width once for both layers?

```{r}
ggplot(mpg, aes(hwy), binwidth = 2) +
  geom_histogram() +
  geom_freqpoly()
```

Nope.

The frequency polygon blends into the histogram, let's make it stand out a bit more:

```{r}
ggplot(mpg, aes(hwy)) +
  geom_histogram(binwidth = 2) +
  geom_freqpoly(binwidth = 2, color = "red")
```

I still hate it. Do the bars need to be solid?

```{r}
ggplot(mpg, aes(hwy)) +
  geom_histogram(binwidth = 2, fill = NA, color = "black", size = 0.2) +
  geom_freqpoly(binwidth = 2, color = "red")
```

## Fuel economy per drivetrain, groupwise

Can I just copy and adapt the code from the last solution?

```{r}
ggplot(mpg, aes(hwy, color = drv)) +
  geom_histogram(fill = NA, color = "black", size = 0.2, binwidth = 2) +
  geom_freqpoly(color = "red", binwidth = 2)
```

Nope. Something's wrong here. Maybe if I follow the template?

```{r}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv),
    binwidth = 2,
    fill = NA,
    color = "black",
    size = 0.2
  )
```

I can't use color both as variable and manual aesthetic, ggplot2 silently uses the manual assignment! Need to be careful with assigning aesthetics.


```{r}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv, group = drv),
    binwidth = 2,
    size = 0.2,
    fill = NA
  ) +
  geom_freqpoly(
    aes(hwy, color = drv, group = drv),
    binwidth = 2
  )
```



```{r}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv),
    binwidth = 2,
    size = 0.2,
    fill = NA
  )
```


```{r}
ggplot(mpg) +
  geom_freqpoly(
    aes(hwy, color = drv),
    binwidth = 2
  )
```

