---
title: "Exercises 1"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

# RStudio walkthrough

1. Evaluate a few arithmetic expressions in the console. Do operator precendence rules hold?

1. Practice:
    - repeating and searching expressions in the history,
    - aborting and restarting the entry of an expression

1. What happens if you type `5 + <enter> 3 <enter>`? Is the result different from `5 <enter> + 3 <enter>`? Why?

1. Practice:
    - restarting the R session
    - other shortcuts from the RStudio cheat sheet

# Getting help

1. Look up the help for the following objects and functions:
    - `diamonds`
    - `geom_boxplot()`
    - `near()`

    Hint: Make sure you have loaded the `tidyverse` package with `library(tidyverse)`.

1. Copy-paste the example for `near()` in your console.

1. Run the example for `geom_boxplot()`.

    Hint: Look up in the help what `example()` does.

1. Find the example for `geom_boxplot()` on http://ggplot2.tidyverse.org or on the USB stick (index.html -> tidyverse -> ggplot2 icon). Do you have a preference for either?

5. The `example()` function is part of base R, can you find it on http://rdocumentation.org? Can you find the other functions you looked up earlier in this example?


# Scatterplots

1. Print the `mpg` dataset to your console.

1. View `mpg` in RStudio's dataset pane (via `View()`). Display all Audis sorted by year then cylinder. What is the maximum number of highway miles per gallon in this dataset?

1. Create a few scatterplots using the following template (here with `hwy` versus `displ`):

    ```{r echo=TRUE}
    ggplot(data = mpg) +
      geom_point(mapping = aes(x = displ, y = hwy))
    ```
    
    Share particularly interesting or boring examples you might encounter.  Did you find anything surprising?

1. Can you plot highway fuel economy given as liters per 100 kilometers against engine displacement?

    Hint: Use the formula `235 / hwy` to convert from miles per gallon.

1. Find more examples in Section 3.2.4 of the "R for data science" book (http://r4ds.had.co.nz or USB stick -> index.html -> "R for data science")


# Scatterplots with three or more variables

1. In the `hwy` vs. `displ` plot, map an additional variable to the "color" aesthetic.  Which cars consume more fuel than expected by the general trend?

1. Experiment with the "color", "shape", "size", and "alpha" aesthetics.  Which combinations of attribute class (categorical/continuous) and aesthetics work well, which don't? Expand on the more surprising examples in the previous exercises.

    Hint: Use `factor(year)` to convert continuous variables with a limited set of values to categorical variables.

1. Can you change both color and shape at the same time? What about the other aesthetics?

1. What happens if you map the same variable to more than one aesthetics?

1. Find more exercises in Section 3.3.1 of r4ds.


# Manual aesthetics

1. Plot `hwy` vs. `displ` with approx. 1/3 opacity for each point, in blue.

    Hint: Use quotes `"` for the color.

1. What happens if you assign a variable, e.g. `year`, to an aesthetic *outside* the `aes()` call?

1. What happens if you assign a constant, e.g. `3`, to the "size" aesthetic *inside* the `aes()` call?

1. What values are valid for `color`, `alpha`, `shape` and `size`?


# Other geoms

1. What do the arguments `se` and `method` to `geom_smooth()` change?

1. What does `geom_rug()` do? Try to reduce overplotting with transparency or by adding `position = "jitter"`. How do you reduce overplotting for the points layer?

1. How does the order of the `geom_...()` calls affect the display?

1. Can you plot both highway and city economy in one plot?

    Hint: The solution to this exercise is probably *not* the recommended way of doing this in ggplot2.  We'll find a better way in a subsequent exercise.

1. Find more exercises in Section 3.6.1 of r4ds.


# Statistical transformations

1. Use a bar plot to find out how many cars of each drivetrain (front/rear/4wd) the `mpg` dataset contains. Which aesthetic mappings do you need to specify? 

    Hint: Find the relevant geom by typing `geom_` on the console or in your script file.

1. Does the appearance of the plot change when you add `y = ..count..` to the `aes()` call? Why/why not?

1. What happens if you instead use `y = ..prop.., group = 1` in the `aes()` call? What happens if you omit `group = 1`? Why?

    Hint: The section "Computed variables" to `geom_bar()` offers a brief explanation.

1. Visualize the overall distribution of fuel economy in the dataset with a histogram.  Compare this with a frequency polygon, use a second layer if you like.

1. Visualize the distribution of fuel economy in the dataset per drivetrain.  Do you prefer a histogram or a frequency polygon?

1. Find more exercises in Section 3.7.1 of r4ds.


# Position adjustments

1. Use a bar chart to show how many cars of each kind of drivetrain the `mpg` dataset contains.

1. What's the most prevalent number of gears for manual or automatic transmissions?

1. Which aesthetic can you map to further discriminate by car class? Which position adjustment is most useful to detect missing combinations of drivetrain and car class?

1. Draw a boxplot of highway fuel economy versus drivetrain.  Is fuel economy also affected by the number of cylinders?

    Hint: Use `factor()` as necessary.

1. Find more exercises in Section 3.8.1 of r4ds.


# Faceting

1. Has fuel economy changed considerably between 1999 and 2008? Perhaps there is a difference if you also consider the car class? Experiment with `facet_wrap()`, `facet_grid()`, aesthetic mappings, and smoothing layers.

1. What changes if you add the argument `labeller = "label_both"` to the `facet_wrap()` call?

1. Is it possible to use a different x and y scale for each facet? How?

1. Experiment with other arguments to `facet_wrap()` and `facet_grid()`.

1. Find more exercises in Section 3.5.1 of r4ds.


# Scales

1. Choose one or two of your previous plots, and give the scales a proper name.

    Hint: Use one or more `scale_...()` functions.

1. What happens if you call the same `scale_...()` function twice in a plot?

1. Add an overview over all available colorblind friendly Brewer scales to your rmarkdown document.

    Hint: Insert a code chunk containing `RColorBrewer::display.brewer.all(colorblindFriendly = TRUE)`

1. Choose an appealing scale for one of your plots that use the "color" or "fill" aesthetics. Apply it with `scale_..._brewer("<name>", palette = "<palette>")`

    Hint: Use `scale_..._distiller()` if you have mapped a continuous variable.

1. Practice Markdown: Create sections, subsections, item lists, emphasized and bold text, ... .


# Coordinate systems

1. Create a flipped boxplot of fuel economy by transmission type.

1. Plot highway vs. city fuel economy with an additional `geom_abline()` layer.  How do you set a 45° degree for the diagonal line? 

1. Try to "estimate" an approximate ratio between the two economy measures by tweaking the `slope` argument to `geom_abline()`.


# Theming

1. Choose your favorite among the predefined themes.

    Hint: All start with `theme_...()`, but watch out for `theme_set()`.

1. Apply `theme(legend.position = "bottom")` to a plot with a color or fill legend.  What happens if you then apply `theme_bw()`? Why?
