---
title: "Manual aesthetics"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Blue transparent points

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    alpha = 1 / 3,
    color = "blue"
  )
```

Some appear darker due to overplotting.


## Assigning variable outside `aes()`

Fails with an error:

```{r echo=TRUE, error=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    color = year
  )
```

The error message is fairly confusing in some cases:

```{r echo=TRUE, error=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    color = class
  )
```


## Assigning constant inside `aes()`

This works, but creates a useless legend and may give unexpected results:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "red")
  )

ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "blue")
  )

ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = "triangle")
  )
```


## Valid values for manual aesthetics

- "color": basic colors, numbers

- "alpha": numbers between 0 and 1

- "shape": integers from 0 to 25, negative integers seem to display a letter

- "size": positive numbers
