---
title: "Scatterplots"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Show mpg on console

```{r}
mpg
```


## RStudio's view pane

```r
View(mpg)
```

- Filtering works by clicking on the "Filter" button, then entering a text (for text columns) or restricting the range of values (for numeric columns).
- Arranging works as usual by clicking on a column header. Because the sort operation is stable, you can sort by two variables by first sorting by the *second* variable, then by the *first*.
- The 1999 VW Jetta 1.9 and VW New Beetle 1.9 have the best highway economy, both will drive you 44 miles for each gallon of fuel.
- Note that the view pane currently shows at most 100 columns.

## Scatterplots

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

More examples:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = drv))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = year, y = cyl))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = manufacturer))
```


## Convert units

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = 235 / hwy))
```
