---
title: "Statistical transformations"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Cars by drivetrain

We need only the "x" aesthetic, "y" is computed automatically.

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv))
```

## Explicit count

The output is the same:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..count..))
```

By default, `geom_bar()` plots the computed "count" aesthetic, which is by convention written with two surrounding dots.


# Proportions

Plotting overall proportions changes the y scale:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..prop.., group = 1))
```

Omitting `group` will plot proportions per drivetrain in each column, which are 1 by default:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..prop..))
```


# Fuel economy

We use a histogram of `hwy`, again only the "x" aesthetic needs to be supplied:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = hwy))
```

The bin width should be set explicitly to avoid the warning:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy),
    binwidth = 1
  )
```

The frequency polygon is very similar. We tweak aesthetics manually in order to be able to distinguish between both layers.

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy),
    fill = NA,
    color = "blue",
    binwidth = 1
  ) +
  geom_freqpoly(
    mapping = aes(x = hwy),
    binwidth = 1,
    size = 2
  )
```


# Fuel economy by drivetrain

We can set the "fill" aesthetic:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy, fill = drv),
    binwidth = 1
  )
```

However, some bars are hidden. This happens less if we use color instead:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy, color = drv),
    fill = NA,
    binwidth = 1
  )
```

The frequency polygon offers a much clearer display of the same data:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_freqpoly(
    mapping = aes(x = hwy, color = drv),
    binwidth = 1
  )
```
