---
title: "My first document"
author: "Kirill Müller"
date: "18 5 2018"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(nycflights13)
library(here)

# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = FALSE)
```

## Heading

```{r message=FALSE}
getwd()

example1_path <- here("data", "example1.csv")

example1 <- read_delim(
  example1_path,
  ";"
)

example1
```


This is some running text.

- Bullet
- Bullet

1. Numbered list
2. Numbered list

*emphasis*

**bold**

### Minor heading

Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.


Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.
Paragraph of text.

```{r}
flights %>% 
  filter(dest == "HNL")
```


```{r}
mpg %>% 
  ggplot() +
  geom_point(aes(x = displ, y = hwy, color = class))
```

