---
title: "Arranging"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)

view <- function(x) print(x)
```

## Flight with shortest airtime

On what day did the flight with the shortest airtime take place?

Hint: Use `head()` to restrict your result to one row only.

```r
flights %>% 
  arrange(___) %>%
  head(1)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r filtering-9}
flights %>% 
  arrange(air_time) %>%
  head(1)
```
</details>

## Flight with heaviest delay

Which flights had the heaviest delays? Can you use the `tail()` verb to obtain this information?

```r
flights %>% 
  arrange(___) %>%
  tail(1)

flights %>% 
  arrange(desc(___)) %>%
  ___(1)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r filtering-10}
flights %>% 
  arrange(arr_delay) %>%
  tail(1)
```

Why doesn't this give the result we're looking for? Can we use a filter?

```{r filtering-11}
flights %>% 
  filter(!is.na(arr_delay)) %>%
  arrange(arr_delay) %>%
  tail(1)
```

Or the pattern below?

```{r filtering-12}
flights %>% 
  arrange(!is.na(arr_delay), arr_delay) %>%
  tail(1)
```

Usually it's easiest to sort in descending order:

```{r filtering-13}
flights %>% 
  arrange(-arr_delay) %>%
  head(1)
```

```{r filtering-14}
flights %>% 
  arrange(desc(arr_delay)) %>%
  head(1)
```
</details>

## Flight with longest airtime

On what day did the flight with the longest airtime take place?

```r
flights %>% 
  arrange(___ - ___) %>%
  head(1)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r filtering-15}
flights %>% 
  arrange(desc(air_time)) %>%
  head(1)
```
</details>

## UA flights with lowest delay

Find two equivalent ways to select the six `"UA"` flights with the lowest delay. Which is faster? Why?

Hint: RStudio has shortcuts for swapping the current line with the next or previous line.

```r
flights %>%
  filter(___) %>%
  arrange(___)

flights %>%
  arrange(___) %>%
  filter(___)
```

<details>
<summary>&#x25BA; Solution:</summary>
If we filter first, fewer observations need to be sorted.

```{r filtering-16}
flights %>% 
  filter(carrier == "UA") %>%
  arrange(arr_delay)
```


```{r filtering-17}
flights %>% 
  arrange(arr_delay) %>%
  filter(carrier == "UA")
```
</details>

## Recovering delay

Which flights were best in recovering from delay in the air?

<details>
<summary>&#x25BA; Solution:</summary>
Recovering from delay means that the arrival delay is lower than the departure delay,
or that arrival minus departure delay is negative.
If we arrange by arrival minus departure delay, negative values are sorted first,
so they are easier to inspect.

```{r filtering-18}
flights %>% 
  arrange(arr_delay - dep_delay)
```
</details>
