---
title: "Arranging"
author: "Kirill Müller, cynkra GmbH"
date: "June 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)

View <- function(x) print(x)
```

## Flight with shortest airtime

On what day did the flight with the shortest airtime take place?

Hint: Use `head()` to restrict your result to one row only.

```r
flights %>% 
  arrange(___) %>%
  head(1)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r filtering-9}
flights %>% 
  arrange(air_time) %>%
  head(1)
```
</details>

## Flight with heaviest delay

Which flights had the heaviest delays? Can you use the `tail()` verb to obtain this information?

```r
flights %>% 
  arrange(___) %>%
  tail(1)

flights %>% 
  arrange(desc(___)) %>%
  ___(1)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r filtering-10}
flights %>% 
  arrange(arr_delay) %>%
  tail(1)
```

Why doesn't this give the result we're looking for? Can we use a filter?

```{r filtering-11}
flights %>% 
  filter(!is.na(arr_delay)) %>%
  arrange(arr_delay) %>%
  tail(1)
```

Or the pattern below?

```{r filtering-12}
flights %>% 
  arrange(!is.na(arr_delay), arr_delay) %>%
  tail(1)
```

Usually it's easiest to sort in descending order:

```{r filtering-13}
flights %>% 
  arrange(-arr_delay) %>%
  head(1)
```

```{r filtering-14}
flights %>% 
  arrange(desc(arr_delay)) %>%
  head(1)
```
</details>

## Flight with longest airtime

On what day did the flight with the longest airtime take place?

```r
flights %>% 
  arrange(___ - ___) %>%
  head(1)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r filtering-15}
flights %>% 
  arrange(desc(air_time)) %>%
  head(1)
```
</details>

## UA flights with lowest delay

Find two equivalent ways to select the six `"UA"` flights with the lowest delay. Which is faster? Why?

Hint: RStudio has shortcuts for swapping the current line with the next or previous line.

```r
flights %>%
  filter(___) %>%
  arrange(___)

flights %>%
  arrange(___) %>%
  filter(___)
```

<details>
<summary>&#x25BA; Solution:</summary>
If we filter first, fewer observations need to be sorted.

```{r filtering-16}
flights %>% 
  filter(carrier == "UA") %>%
  arrange(arr_delay)
```


```{r filtering-17}
flights %>% 
  arrange(arr_delay) %>%
  filter(carrier == "UA")
```
</details>

## Recovering delay

Which flights were best in recovering from delay in the air?

TODO

<details>
<summary>&#x25BA; Solution:</summary>
```{r filtering-18}
flights %>% 
  arrange(dep_delay - arr_delay)
```
</details>
