---
title: "Coordinate systems"
author: "Kirill Müller, cynkra GmbH"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(20180516)
knitr::opts_chunk$set(cache = TRUE)
```

## Flipped boxplot

Create a flipped boxplot of fuel economy by transmission type.

```{r coord-1, echo=TRUE}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = trans, y = hwy)) +
  coord_flip()
```

## Compare highway and city

Plot highway vs. city fuel economy with an additional `geom_abline()` layer.  How do you set a 45° degree for the diagonal line? 

```{r coord-2, echo=TRUE}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = cty, y = hwy)) +
  geom_abline(color = "red")
```

The `coord_fixed()` calls fixes the aspect ratio, and also the angle of the red line with slope 1:

```{r coord-3, echo=TRUE}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = cty, y = hwy)) +
  geom_abline(color = "red") +
  coord_fixed()
```


## Tweaking the slope

Try to "estimate" an approximate ratio between the two economy measures by tweaking the `slope` argument to `geom_abline()`.

Guessing 40 / 30 from the previous plot:

```{r coord-4, echo=TRUE}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = cty, y = hwy)) +
  geom_abline(slope = 40 / 30, color = "red") +
  coord_fixed()
```
