---
title: "Facets"
author: "Kirill Müller, cynkra GmbH"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Improvement over time?

Has fuel economy changed considerably between 1999 and 2008? Perhaps there is a difference if you also consider the car class? Experiment with `facet_wrap()`, `facet_grid()`, aesthetic mappings, and smoothing layers.

```
ggplot(_____) +
  geom_point() +
  facet_wrap(~___)
```

```
ggplot(_____) +
  geom_point() +
  geom_smooth() + 
  facet_grid(___ ~ ___)
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
We can split the simple scatterplot over two facets, one per year:

```{r facet-1, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~year)
```

The smoothing layer seems to confirm a slight improvement, especially for engines with a displacement of three or more liters:

```{r facet-2, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~year)
```

We can also look at each car class in parallel.

```{r facet-3, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(year~class)
```

But too many facets may be not as helpful, we can also use color:

```{r facet-4, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = factor(year))
  ) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = factor(year)),
    method = "lm"
  ) +
  facet_wrap(~class)
```
</details>


## Labeling the facets

What changes if you add the argument `labeller = "label_both"` to the `facet_wrap()` call?

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
This is useful if the value does not speak for itself:

```{r facet-5, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~cyl, labeller = "label_both")
```
</details>


## Different scales

Is it possible to use a different x and y scale for each facet? How?

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
Via the `scales` argument, zooms in to the range of the corresponding scale(s).

```{r facet-6, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(
    ~cyl,
    labeller = "label_both",
    scales = "free_x"
  )
```

```{r facet-7, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(
    ~cyl,
    labeller = "label_both",
    scales = "free_y"
  )
```

```{r facet-8, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(
    ~cyl,
    labeller = "label_both",
    scales = "free"
  )
```
<<<<<<< HEAD
</details>

## Other arguments

Experiment with other arguments to `facet_wrap()` and `facet_grid()`.

## More exercises
=======

##

Experiment with other arguments to `facet_wrap()` and `facet_grid()`.

##
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

Find more exercises in Section 3.5.1 of r4ds.
