---
title: "Facets"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Improvement over time?

We can split the simple scatterplot over two facets, one per year:

```{r facet-1, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~year)
```

The smoothing layer seems to confirm a slight improvement, especially for engines with a displacement of three or more liters:

```{r facet-2, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~year)
```

We can also look at each car class in parallel.

```{r facet-3, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(year~class)
```

But too many facets may be not as helpful, we can also use color:

```{r facet-4, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = factor(year))
  ) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = factor(year)),
    method = "lm"
  ) +
  facet_wrap(~class)
```


## Labeling the facets

This is useful if the value does not speak for itself:

```{r facet-5, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~cyl, labeller = "label_both")
```


## Different scales

Via the `scales` argument, zooms in to the range of the corresponding scale(s).

```{r facet-6, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(
    ~cyl,
    labeller = "label_both",
    scales = "free_x"
  )
```

```{r facet-7, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(
    ~cyl,
    labeller = "label_both",
    scales = "free_y"
  )
```

```{r facet-8, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(
    ~cyl,
    labeller = "label_both",
    scales = "free"
  )
```
