---
title: "Filtering and plotting"
author: "Kirill Müller, cynkra GmbH"
date: "Nov 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)
```

## Histogram of air time of all flights

Plot a histogram of the air time of all flights. Exclude Honolulu International Airport in Hawaii to get rid of the peak at the right-hand side. Zoom into the flights that have an air time between 400 and 500 minutes.

Hint: Start with `flights %>% ggplot() + ...`

```r
flights %>% 
  ggplot(___) +
    ___()

flights %>% 
  filter(___) %>%
  ggplot(___) +
    ___()

flights %>% 
  filter(___) %>%
  filter(___) %>%
  ___
```

```{r filtering-23}
flights %>%
  ggplot() +
  geom_histogram(
    aes(x = air_time),
    na.rm = TRUE,
    binwidth = 15
  )
```

```{r filtering-24}
flights %>%
  filter(dest != "HNL") %>%
  ggplot() +
  geom_histogram(
    aes(x = air_time),
    na.rm = TRUE,
    binwidth = 15
  )
```

```{r filtering-25}
flights %>%
  filter(dest != "HNL") %>%
  filter(between(air_time, 400, 500)) %>%
  ggplot() +
  geom_histogram(
    aes(x = air_time),
    na.rm = TRUE,
    binwidth = 10
  )
```

## All very close relations

Plot a heat map for all relations with an air time shorter than one hour.

Hint: Use `geom_bin2d()`.

```r
flights %>% 
  filter(___) %>%
  ggplot(___) +
    ___()
```

```{r filtering-26}
flights %>%
  filter(air_time < 60) %>%
  ggplot() +
  geom_bin2d(aes(origin, dest))
```

##

Think of other plots of the `flights` data that would not work if applied on the full dataset but are useful when applying a filter beforehand.
