---
title: "Filtering"
author: "Kirill Müller, cynkra GmbH"
date: "June 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)

View <- function(x) print(x)
```

## Flights on the same date as today

Find all flights that departed today x years ago, flown by `"US"`. Two simple solutions exist, which appeals most to you?

```r
flights %>%
  filter(___, ___) %>%
  filter(___)
  
flights %>%
  filter(_____)
```  

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r filter-combine-1}
flights %>%
  filter(day == 13, month == 11) %>%
  filter(carrier == "US")
  
flights %>%
  filter(day == 13 & month == 11 & carrier == "US")
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## Red eye flights

Find all flights that departed before 6:00 AM or after 10:00 PM.

```r
flights %>%
  filter(___ | ___)
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r filter-combine-2}
flights %>%
  filter(dep_time < 600 | dep_time > 2200)
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## Flights not by UA or WN

Find all flights not flown by either `"UA"` or `"WN"`. Can you think of three different solutions? Which appeals more to you?

```r
flights %>%
  filter(___ ___ ___)

flights %>%
  filter(!(___) ___ !(___))

flights %>%
  filter(!(_____))
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r filter-combine-3}
flights %>%
  filter(carrier != "UA" & carrier != "WN")

flights %>%
  filter(!(carrier == "UA" | carrier == "WN"))

flights %>%
  filter(!(carrier %in% c("UA", "WN")))
```
</details>

## Missing data

Which flights have a missing departure or arrival time? Which have both missing? Can the number of flights that have a missing arrival but not departure time correspond to lost or crashed flights?

```r
flights %>%
  filter(is.na(___))

flights %>%
  filter(___(___) ___ _____)

flights %>%
  filter(_____ ___ !_____)
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r filter-combine-4}
flights %>%
  filter(is.na(arr_time))

flights %>%
  filter(is.na(arr_time) & is.na(dep_time))

flights %>%
  filter(is.na(arr_time) & !is.na(dep_time))
```
</details>

## Commutativity

Find two equivalent ways to select the six `"UA"` flights with the lowest delay. Which is faster? Why?

Hint: RStudio has shortcuts for swapping the current line with the next or previous line.

```r
flights %>%
  filter(___) %>%
  arrange(___)

flights %>%
  arrange(___) %>%
  filter(___)
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
Arranging a smaller subset needs less work.

```{r filter-combine-5}
system.time(
  flights %>%
    arrange(arr_delay) %>% 
    filter(carrier == "UA")
)

system.time(
  flights %>%
    filter(carrier == "UA") %>%
    arrange(arr_delay)
)
```
<<<<<<< HEAD
</details>

## Understand predicate logic

Look at the "Details" section in the help page for `|` with `help("|")` to understand predicate logic in R. (We need element-wise comparisons.)

## More exercises
=======

##

Look at the "Details" section in the help page for `|` with `help("|")` to understand predicate logic in R. (We need element-wise comparisons.)

##
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

Find more exercises in items 1 and 4 of Section 5.2.4, and in Section 5.3.1, of r4ds.
