---
title: "Filtering"
author: "Kirill Müller"
date: "June 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)

View <- function(x) print(x)
```

## All flights on this day x years ago

Be careful with the equality operator `==`:

```{r, error = TRUE}
flights %>% 
  filter(month = 6, day = 2)

flights %>% 
  filter(month == 6, day == 2)
```

Can we make a dynamic query?

```{r}
flights %>% 
  filter(
    month == lubridate::month(Sys.Date()),
    day == lubridate::day(Sys.Date())
  )
```

Can we use arguments?

```{r}
month_ <- lubridate::month(Sys.Date())
day_ <- lubridate::day(Sys.Date())
flights %>% 
  filter(
    month == month_,
    day == day_
  )
```


## All flights between 8:00 AM and 10:00 AM

```{r}
flights %>% 
  filter(dep_time >= 800, dep_time <= 2200)
```

```{r}
flights %>% 
  filter(between(dep_time, 800, 2200))
```

## Flights in winter months

```{r}
flights %>%
  filter(month %in% c(12, 1, 2))
```

```{r}
winter_months <- c(12, 1, 2)
flights %>%
  filter(month %in% winter_months)
```

## Departure time later than arrival time

```{r}
flights %>% 
  filter(dep_time > arr_time)
```

## Flight with shortest airtime

```{r}
flights %>% 
  arrange(air_time) %>%
  head(1)
```


## Flight with heaviest delay

```{r}
flights %>% 
  arrange(arr_delay) %>%
  tail(1)
```

Why doesn't this give the result we're looking for? Can we use a filter?

```{r}
flights %>% 
  filter(!is.na(arr_delay)) %>%
  arrange(arr_delay) %>%
  tail(1)
```

Or the pattern below?

```{r}
flights %>% 
  arrange(!is.na(arr_delay), arr_delay) %>%
  tail(1)
```

Usually it's easiest to sort in descending order:

```{r}
flights %>% 
  arrange(-arr_delay) %>%
  head(1)
```

```{r}
flights %>% 
  arrange(desc(arr_delay)) %>%
  head(1)
```


## Flight with longest airtime

```{r}
flights %>% 
  arrange(desc(air_time)) %>%
  head(1)
```


## UA flights with lowest delay

If we filter first, fewer observations need to be sorted.

```{r}
flights %>% 
  filter(carrier == "UA") %>%
  arrange(arr_delay)
```


```{r}
flights %>% 
  arrange(arr_delay) %>%
  filter(carrier == "UA")
```


## Recovering delay

```{r}
flights %>% 
  arrange(dep_delay - arr_delay)
```

## Intermediate variables

Naming is hard!

```{r}
early_flights <- filter(flights, dep_time >= 600)
early_late_flights <-
  filter(early_flights, arr_time >= 2200)
early_late_ua_flights <-
  filter(early_late_flights, carrier == "UA")
early_late_late_ua_flights <-
  filter(early_late_ua_flights, arr_delay > 120)
early_late_late_ua_flights_not_honolulu <-
  filter(early_late_late_ua_flights, dest != "HNL")
early_late_late_ua_flights_not_honolulu_sorted <-
  arrange(
    early_late_late_ua_flights_not_honolulu,
    distance
  )
View(early_late_late_ua_flights_not_honolulu_sorted)
```

## Nested expressions

Difficult to read.

```{r}
View(
  arrange(
    filter(
      filter(
        filter(
          filter(
            filter(
              flights,
              dep_time <= 600
            ),
            arr_time >= 2200
          ),
          carrier == "UA"
        ),
        arr_delay > 120
      ),
      dest != "HNL"
    ),
    distance
  )
)
```

## Pipe

```{r}
flights %>% 
  filter(dep_time <= 600) %>% 
  filter(arr_time >= 2200) %>% 
  filter(carrier == "UA") %>% 
  filter(arr_delay > 120) %>% 
  filter(dest != "HNL") %>%
  arrange(distance) %>%
  View()
```

The original data is never updated! You still need to assign the result of a pipe to a variable:

```{r}
late_late_ua_flights_not_honolulu <-
  flights %>% 
  filter(dep_time <= 600) %>% 
  filter(arr_time >= 2200) %>% 
  filter(carrier == "UA") %>% 
  filter(arr_delay > 120) %>% 
  filter(dest != "HNL") %>%
  arrange(distance)
```

## Histogram of air time of all flights

```{r}
flights %>% 
  ggplot() +
  geom_histogram(
    aes(x = air_time),
    na.rm = TRUE,
    binwidth = 15
  )
```


```{r}
flights %>%
  filter(dest != "HNL") %>%
  ggplot() +
  geom_histogram(
    aes(x = air_time),
    na.rm = TRUE,
    binwidth = 15
  )
```


```{r}
flights %>%
  filter(dest != "HNL") %>%
  filter(between(air_time, 400, 500)) %>% 
  ggplot() +
  geom_histogram(
    aes(x = air_time),
    na.rm = TRUE,
    binwidth = 10
  )
```

## All very close relations

```{r}
flights %>%
  filter(air_time < 60) %>% 
  ggplot() +
  geom_bin2d(aes(origin, dest))
```
