---
title: "Other geoms"
author: "Kirill Müller, cynkra GmbH"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Arguments to `geom_smooth()`

Try `geom_smooth()`. What do the arguments `se` and `method` to `geom_smooth()` change?

```    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy),
    se = ___,
    method = ___
  )
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
`method` uses a different model to fit the data:

```{r geoms-1, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy), method = "lm")
```

`se = FALSE` turns off the confidence band:

```{r geoms-2, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy), se = FALSE)
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## The rug

What does `geom_rug()` do? Try to reduce overplotting with transparency or by adding `position = "jitter"`. How do you reduce overplotting for the points layer?

```    
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    ___ = ___
  ) +
  geom_rug(
    mapping = aes(x = displ, y = hwy),
    ___ = ___
  )
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
Plots marginal distributions of the data close to the axes.

```{r geoms-3, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_rug(mapping = aes(x = displ, y = hwy))
```

To reduce overplotting, the "alpha" aesthetic can be set independently for each geom to a constant value:

```{r geoms-4, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    alpha = 0.3
  ) +
  geom_rug(
    mapping = aes(x = displ, y = hwy)
  )
```

```{r geoms-5, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy)
  ) +
  geom_rug(
    mapping = aes(x = displ, y = hwy),
    alpha = 0.3
  )
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## Order of `geom_...()` calls

How does the order of the `geom_...()` calls affect the display?

```
ggplot(data = ___, mapping = aes(_____)) +
  geom_point() +
  geom_smooth()
```
    
```
ggplot(data = ___, mapping = aes(_____)) +
  geom_smooth() +
  geom_point()
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
The geoms are painted in order of appearance:

```{r geoms-6, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r geoms-7, echo=TRUE}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## Compare highway and city

Can you plot both highway and city economy in one plot?

Hint: The solution to this exercise is *not* the recommended way of doing this in ggplot2.  We'll find a better way in a subsequent exercise.

```
ggplot(_____) +
  geom_point(mapping = _____, color = "___") +
  geom_point(mapping = _____, color = "___")
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
We *could* add two layers, each with a different color. But this still doesn't give us a legend.

```{r geoms-8, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "red") +
  geom_point(mapping = aes(x = displ, y = cty), color = "blue")
```

We need better data transformation tools to reformat the data for plotting it in a more natural way.
</details>

## Cars by drivetrain

Use a bar plot to find out how many cars of each drivetrain (front/rear/4wd) the `mpg` dataset contains. Which aesthetic mappings do you need to specify? 

Hint: Find the relevant geom by typing `geom_` on the console or in your script file.

```    
ggplot(_____, aes(_____)) +
  geom_bar()
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
I tried `geom_histogram()` and `geom_col()`, neither worked. The histogram is for continuous data only, for `geom_col()` I'd need to supply actual counts which I don't have. The `geom_bar()` function computes the counts for me by applying the `"count"` statistical transformation to my data before plotting.

We need only the "x" aesthetic, "y" is computed automatically. `drv` is the relevant variable.

```{r geoms-9}
ggplot(mpg) +
  geom_bar(aes(x = drv))
```
</details>

## Explicit count

Does the appearance of the plot change when you add `y = ..count..` to the `aes()` call? Why/why not?

```    
ggplot(_____, aes(_____, y = ..count..)) +
  geom_bar()
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
The output is the same:

```{r geoms-10, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..count..))
```

By default, `geom_bar()` plots the computed "count" aesthetic, which is by convention written with two surrounding dots.
</details>


## Proportions

What happens if you instead use `y = ..prop.., group = 1` in the `aes()` call? What happens if you omit `group = 1`? Why?

Hint: The section "Computed variables" to `geom_bar()` offers a brief explanation.

```
ggplot(_____, aes(_____, y = ..prop.., group = 1)) +
  geom_bar()
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
Plotting overall proportions changes the y scale:

```{r geoms-11, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..prop.., group = 1))
```

Omitting `group` will plot proportions per drivetrain in each column, which are 1 by default:

```{r geoms-12, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..prop..))
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## Fuel economy

Visualize the overall distribution of fuel economy in the dataset with a histogram.  Compare this with a frequency polygon, use a second layer if you like.
<<<<<<< HEAD
=======

```
ggplot(data = mpg, mapping = aes(x = hwy)) +
  geom_____()
```

We use a histogram of `hwy`, again only the "x" aesthetic needs to be supplied:
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

```
ggplot(data = mpg, mapping = aes(x = hwy)) +
  geom_____()
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
We use a histogram of `hwy`, again only the "x" aesthetic needs to be supplied:
=======
How do you remove the warning in the previous example?
    
```
ggplot(_____) +
  geom_____(binwidth = ___)
```

The bin width should be set explicitly to avoid the warning:
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

```{r geoms-13, echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = hwy))
```

The frequency polygon is very similar. We tweak aesthetics manually in order to be able to distinguish between both layers.

```{r geoms-15, echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy),
    fill = NA,
    color = "blue",
    binwidth = 1
  ) +
  geom_freqpoly(
    mapping = aes(x = hwy),
    binwidth = 1,
    size = 2
  )
```
</details>

## Silencing a warning

How do you remove the warning in the previous example?
    
```
ggplot(_____) +
  geom_____(binwidth = ___)
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
The bin width should be set explicitly to avoid the warning:

```{r geoms-14, echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy),
    binwidth = 1
  )
```
</details>

=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
## Fuel economy by drivetrain

Visualize the distribution of fuel economy in the dataset per drivetrain.  Do you prefer a histogram or a frequency polygon?

<details>
<summary>&#x25BA; Solution:</summary>
We can set the "fill" aesthetic:

```{r geoms-16, echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy, fill = drv),
    binwidth = 1
  )
```

However, the bars are stacked, and difficult to compare to each other:

```{r geoms-17, echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy, color = drv),
    fill = NA,
    binwidth = 1
  )
```

The frequency polygon offers a much clearer display of the same data:

```{r geoms-18, echo=TRUE}
ggplot(data = mpg) +
  geom_freqpoly(
    mapping = aes(x = hwy, color = drv),
    binwidth = 1
  )
```

### A history of failed attempts

Can I just copy and adapt the code from the last solution?

```{r geoms-19}
ggplot(mpg, aes(hwy, color = drv)) +
  geom_histogram(fill = NA, color = "black", size = 0.2, binwidth = 2) +
  geom_freqpoly(color = "red", binwidth = 2)
```

Nope. Something's wrong here. Maybe if I follow the template?

```{r geoms-20}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv),
    binwidth = 2,
    fill = NA,
    color = "black",
    size = 0.2
  )
```

I can't use color both as variable and manual aesthetic, ggplot2 silently uses the manual assignment! Need to be careful with assigning aesthetics.

```{r geoms-21}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv, group = drv),
    binwidth = 2,
    size = 0.2,
    fill = NA
  ) +
  geom_freqpoly(
    aes(hwy, color = drv, group = drv),
    binwidth = 2
  )
```

```{r geoms-22}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv),
    binwidth = 2,
    size = 0.2,
    fill = NA
  )
```

```{r geoms-23}
ggplot(mpg) +
  geom_freqpoly(
    aes(hwy, color = drv),
    binwidth = 2
  )
```
<<<<<<< HEAD
</details>

## More exercises
=======

## 

Visualize the distribution of fuel economy in the dataset per drivetrain.  Do you prefer a histogram or a frequency polygon?

##
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

Find more exercises in Sections 3.6.1 and 3.7.1 of r4ds.
