---
title: "Geoms"
author: "Kirill Müller, Patrick Schratz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Arguments to `geom_smooth()`

Try `geom_smooth()`. What do the arguments `se` and `method` to `geom_smooth()` change?

```    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy),
    se = ___,
    method = ___
  )
```

<details>
<summary>&#x25BA; Solution:</summary>
`method` uses a different model to fit the data:

```{r geoms-1, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy), method = "lm")
```

`se = FALSE` turns off the confidence band:

```{r geoms-2, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy), se = FALSE)
```
</details>

## The rug

What does `geom_rug()` do? Try to reduce overplotting with transparency or by adding `position = "jitter"`. How do you reduce overplotting for the points layer?

```    
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    ___ = ___
  ) +
  geom_rug(
    mapping = aes(x = displ, y = hwy),
    ___ = ___
  )
```

<details>
<summary>&#x25BA; Solution:</summary>
Plots marginal distributions of the data close to the axes.

```{r geoms-3, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_rug(mapping = aes(x = displ, y = hwy))
```

To reduce overplotting, the "alpha" aesthetic can be set independently for each geom to a constant value:

```{r geoms-4, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    alpha = 0.3
  ) +
  geom_rug(
    mapping = aes(x = displ, y = hwy)
  )
```

```{r geoms-5, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy)
  ) +
  geom_rug(
    mapping = aes(x = displ, y = hwy),
    alpha = 0.3
  )
```
</details>

## Order of `geom_...()` calls

How does the order of the `geom_...()` calls affect the display?

```
ggplot(data = ___, mapping = aes(_____)) +
  geom_point() +
  geom_smooth()
```
    
```
ggplot(data = ___, mapping = aes(_____)) +
  geom_smooth() +
  geom_point()
```

<details>
<summary>&#x25BA; Solution:</summary>
The geoms are painted in order of appearance:

```{r geoms-6, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r geoms-7, echo=TRUE}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
</details>

## Compare highway and city

Can you plot both highway and city economy in one plot?

Hint: The solution to this exercise is *not* the recommended way of doing this in ggplot2.  We'll find a better way in a subsequent exercise.

```
ggplot(_____) +
  geom_point(mapping = _____, color = "___") +
  geom_point(mapping = _____, color = "___")
```

<details>
<summary>&#x25BA; Solution:</summary>
We *could* add two layers, each with a different color. But this still doesn't give us a legend.

```{r geoms-8, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "red") +
  geom_point(mapping = aes(x = displ, y = cty), color = "blue")
```

We need better data transformation tools to reformat the data for plotting it in a more natural way.
</details>

## Cars by drivetrain

Use a bar plot to find out how many cars of each drivetrain (front/rear/4wd) the `mpg` dataset contains. Which aesthetic mappings do you need to specify? 

Hint: Find the relevant geom by typing `geom_` on the console or in your script file.

```    
ggplot(_____, aes(_____)) +
  geom____()
```

<details>
<summary>&#x25BA; Solution:</summary>
I tried `geom_histogram()` and `geom_col()`, neither worked. The histogram is for continuous data only, for `geom_col()` I'd need to supply actual counts which I don't have. The `geom_bar()` function computes the counts for me by applying the `"count"` statistical transformation to my data before plotting.

We need only the "x" aesthetic, "y" is computed automatically. `drv` is the relevant variable.

```{r geoms-9}
ggplot(mpg) +
  geom_bar(aes(x = drv))
```
</details>

## More exercises

Find more exercises in Sections 3.6.1 and 3.7.1 of r4ds.
