---
title: "ggplot2 exercises part 1"
author: "Kirill Müller"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## [Scatterplots](scatter.html)

1. Print the `mpg` dataset to your console.

1. View `mpg` in RStudio's dataset pane (via `View()`). Display all Audis sorted by year then cylinder. What is the maximum number of highway miles per gallon in this dataset?

1. Create a few scatterplots using the following template (here with `hwy` versus `displ`):

    ```{r echo=TRUE, eval=FALSE}
    ggplot(data = mpg) +
      geom_point(mapping = aes(x = displ, y = hwy))
    ```
    
    Share particularly interesting or boring examples you might encounter.  Did you find anything surprising?

1. Can you plot highway fuel economy given as liters per 100 kilometers against engine displacement?

    Hint: Use the formula `235 / hwy` to convert from miles per gallon.
    
1. How can you reduce overplotting? Use your favorite search engine to find out.

1. Find more examples in Section 3.2.4 of the "R for data science" book (http://r4ds.had.co.nz).


## [Scatterplots with three or more variables](scatter3.html)

1. In the `hwy` vs. `displ` plot, map an additional variable to the "color" aesthetic.  Which cars consume more fuel than expected by the general trend?

1. Experiment with the "color", "shape", "size", and "alpha" aesthetics.  Which combinations of attribute class (categorical/continuous) and aesthetics work well, which don't? Expand on the more surprising examples in the previous exercises.

    Hint: Use `factor(year)` to convert continuous variables with a limited set of values to categorical variables.

1. Can you change both color and shape at the same time? What about the other aesthetics?

1. What happens if you map the same variable to more than one aesthetic?

1. Find more exercises in Section 3.3.1 of r4ds.


## [Manual aesthetics](manual.html)

1. Plot `hwy` vs. `displ` with approx. 1/3 opacity for each point, in blue.

    Hint: Use quotes `"` for the color.

1. What happens if you assign a variable, e.g. `year`, to an aesthetic *outside* the `aes()` call?

1. What happens if you assign a constant, e.g. `3`, to the "size" aesthetic *inside* the `aes()` call?

1. What values are valid for `color`, `alpha`, `shape` and `size`?

1. Find how to update the default point color for all scatterplots in the help for `update_geom_defaults()`. Does the setting persist between plots? Is it still active after restarting RStudio (Session → Restart R or Ctrl + Shift + F10, Cmd + Shift + F10 on the Mac).

1. Find how to update the default point color for all label texts in the help for `theme()`. How to set or update the theme for all subsequent plots?

    ```r
    ggplot(_____) +
      _____ +
      theme(___)
    ```

## [Other point geoms](geoms.html)

1. What do the arguments `se` and `method` to `geom_smooth()` change?

1. What does `geom_rug()` do? Try to reduce overplotting with transparency or by adding `position = "jitter"`. How do you reduce overplotting for the points layer?

1. How does the order of the `geom_...()` calls affect the display?

1. Can you plot both highway and city economy in one plot?

    Hint: The solution to this exercise is *not* the recommended way of doing this in ggplot2.  We'll find a better way in a subsequent exercise.

1. Find more exercises in Section 3.6.1 of r4ds.


## [Statistical transformations](stat.html)

1. Use a bar plot to find out how many cars of each drivetrain (front/rear/4wd) the `mpg` dataset contains. Which aesthetic mappings do you need to specify? 

    Hint: Find the relevant geom by typing `geom_` on the console or in your script file.

1. Does the appearance of the plot change when you add `y = ..count..` to the `aes()` call? Why/why not?

1. What happens if you instead use `y = ..prop.., group = 1` in the `aes()` call? What happens if you omit `group = 1`? Why?

    Hint: The section "Computed variables" to `geom_bar()` offers a brief explanation.

1. Visualize the overall distribution of fuel economy in the dataset with a histogram.  Compare this with a frequency polygon, use a second layer if you like.

1. Visualize the distribution of fuel economy in the dataset per drivetrain.  Do you prefer a histogram or a frequency polygon?

1. Find more exercises in Section 3.7.1 of r4ds.


## [Position adjustments](position.html)

1. What's the most prevalent number of gears for manual or automatic transmissions?

1. Which aesthetic can you map to further discriminate by car class? Which position adjustment is most useful to detect missing combinations of drivetrain and car class?

1. Draw a boxplot of highway fuel economy versus drivetrain.  Is fuel economy also affected by the number of cylinders?

    Hint: Use `factor()` as necessary.

1. Find more exercises in Section 3.8.1 of r4ds.


## [Faceting](facet.html)

1. Has fuel economy changed considerably between 1999 and 2008? Perhaps there is a difference if you also consider the car class? Experiment with `facet_wrap()`, `facet_grid()`, aesthetic mappings, and smoothing layers.

1. What changes if you add the argument `labeller = "label_both"` to the `facet_wrap()` call?

1. Is it possible to use a different x and y scale for each facet? How?

1. Experiment with other arguments to `facet_wrap()` and `facet_grid()`.

1. Find more exercises in Section 3.5.1 of r4ds.
