---
title: "Manual aesthetics"
author: "Kirill Müller, cynkra GmbH"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Blue transparent points

Plot `hwy` vs. `displ` with approx. 1/3 opacity for each point, in blue.

Hint: Use quotes `"` for the color.

```    
ggplot(_____) +
  geom_point(
    mapping = aes(_____),
    alpha = ___,
    color = "___"
  )
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r manual-1, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    alpha = 1 / 3,
    color = "blue"
  )
```

Some appear darker due to overplotting.
</details>


## Assigning variable outside `aes()`

What happens if you assign a variable, e.g. `year`, to an aesthetic *outside* the `aes()` call?

```    
ggplot(_____) +
  geom_point(
    mapping = aes(_____),
    color = year
  )
```

<details>
<summary>&#x25BA; Solution:</summary>
Fails with an error:

```{r manual-2, echo=TRUE, error=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    color = year
  )
```

The error message is fairly confusing in some cases:

```{r manual-3, echo=TRUE, error=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    color = class
  )
```
</details>

## Assigning constant inside `aes()`

What happens if you assign a constant, e.g. `3`, to the "size" aesthetic *inside* the `aes()` call?

```    
ggplot(_____) +
  geom_point(
    mapping = aes(_____, size = 3)
  )
```

<details>
<summary>&#x25BA; Solution:</summary>
This works, but creates a useless legend and may give unexpected results:

```{r manual-4, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "red")
  )

ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "blue")
  )

ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = "triangle")
  )
```
</details>

## Valid values for manual aesthetics

What values are valid for `color`, `alpha`, `shape` and `size`?

```    
ggplot(_____) +
  geom_point(
    mapping = aes(_____),
    ___ = ___
  )
```

<details>
<summary>&#x25BA; Solution:</summary>
- "color": basic colors, numbers

- "alpha": numbers between 0 and 1

- "shape": integers from 0 to 25, negative integers seem to display a letter

- "size": positive numbers
</details>

## Default point color

Find how to update the default point color for all scatterplots in the help for `update_geom_defaults()`. Does the setting persist between plots? Is it still active after restarting RStudio (Session → Restart R or Ctrl + Shift + F10, Cmd + Shift + F10 on the Mac).

## Default text color

Find how to update the default color for all label texts in the help for `theme()`. How to set or update the theme for all subsequent plots?

```
ggplot(_____) +
  _____ +
  theme(___)
```
