---
title: "Mutate"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)
```

## Speed as miles per hour

Store the speed for each flight as miles per hour in a new variable.

```r
flights %>% 
  mutate(miles_per_hour = air_time ___ distance ___ ___)
```
 
<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-1}
flights %>%
  mutate(miles_per_hour = distance / air_time * 60)
```
</details>

## Speed as miles per minute

Can you use an intermediate variable to clarify the intent? How do you remove the intermediate variable?

```r
flights %>% 
  mutate(miles_per_minute = _____) %>% 
  mutate(miles_per_hour = _____) %>% 
  select(_____)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-1-1}
flights %>%
  mutate(miles_per_minute = distance / air_time) %>% 
  mutate(miles_per_hour = miles_per_minute * 60) %>% 
  select(-miles_per_minute)
```
</details>

## Speed, visualized

Visualize the speed distribution as a histogram. Would this visualization work without involving `mutate()`?

```r
flights %>% 
  ______ %>%
  ggplot(aes(___)) +
  _____

# Alternative:
flights %>%
  ggplot(aes(___)) +
  _____
```


<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-1-2}
flights %>%
  mutate(miles_per_hour = distance / air_time * 60) %>%
  ggplot() +
  geom_histogram(
    aes(miles_per_hour),
    na.rm = TRUE,
    binwidth = 20
  )
```

```{r mutate-1-3}
flights %>% 
  ggplot() +
  geom_histogram(aes(distance / air_time))
```
</details>

## On time status

Create a new logical variable that indicates if the flight arrived on time.

```r
flights %>%
  mutate(on_time = (_____))
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-1-bis}
flights %>%
  mutate(
    on_time = (arr_delay <= 0)
  )
```
</details>

## On time status, visualized

Visualize the aggregated on-time status per airline with a useful text.

```r
flights %>%
  flights %>%
  mutate(
    on_time = _____,
    on_time_desc = if_else(___, "On time", ___)
  ) %>%
  ggplot(aes(___)) +
  geom_bar()
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-1-bis-1}
flights %>%
  mutate(
    on_time = (arr_delay <= 0),
    on_time_desc = if_else(on_time, "On time", "Delayed")
  ) %>%
  ggplot(aes(x = carrier, fill = on_time_desc)) +
  geom_bar()
```
</details>

## Speed distributions

Can you detect a difference in the speed distributions of on-time vs. delayed flights? Ose color of faceting.

```r
speed_and_on_time_info <-
  _____

speed_and_on_time_info %>%
  ggplot() +
  geom_freqpoly(
    aes(x = ___, y = ..density.., color = ___),
    na.rm = TRUE,
    binwidth = 20
  )

speed_and_on_time_info %>%
  ggplot() +
  geom_histogram(
    aes(x = ___),
    na.rm = TRUE,
    binwidth = 20
  ) +
  facet_wrap(~___, ncol = 1)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-3}
speed_and_on_time_info <-
  flights %>%
  mutate(
    miles_per_minute = distance / air_time,
    miles_per_hour = miles_per_minute * 60
  ) %>% 
  select(-miles_per_minute) %>% 
  mutate(
    on_time = (arr_delay <= 0),
    on_time_desc = if_else(on_time, "On time", "Delayed")
  ) %>% 
  select(-on_time)

speed_and_on_time_info %>% 
  ggplot() +
  geom_freqpoly(
    aes(x = miles_per_hour, y = ..density.., color = on_time_desc),
    na.rm = TRUE,
    binwidth = 20
  )

speed_and_on_time_info %>% 
  filter(!is.na(on_time_desc)) %>% 
  ggplot() +
  geom_histogram(
    aes(x = miles_per_hour),
    na.rm = TRUE,
    binwidth = 20
  ) +
  facet_wrap(~on_time_desc, ncol = 1)
```
</details>

## Date

Create two new variables `date_hour` and `date_ymd`, using `as.Date()` or `lubridate::make_date()`, respectively. Are the two values the same for all observations? What happens if we omit the `tz` argument to `as.Date()`?

```r
flights %>%
  mutate(
    ___ = as.Date(___, tz = "EST"),
    ___ = lubridate::make_date(_____)
  ) %>% 
  filter(___)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-2}
flights %>%
  mutate(
    date_hour = as.Date(time_hour, tz = "EST"),
    date_ymd = lubridate::make_date(year, month, day)
  ) %>% 
  filter(date_hour != date_ymd)
```
</details>

## Deviation from average departure delay

Visualize the deviation from the overall average departure delay for the three airports of New York City. Consider using a violin plot.

```r
flights %>%
  mutate(dep_delay_dev = ___ - mean(___)) %>%
  ggplot(aes(___)) +
  _____ +
  _____
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r mutate-5}
flights %>% 
  mutate(dev = dep_delay - mean(dep_delay, na.rm = TRUE)) %>%
  ggplot() +
  geom_violin(aes(x = origin, y = dev), na.rm = TRUE)
```
</details>

## More exercises

Find more exercises in Section 5.5.2 of r4ds.
