---
title: "Data visualisation, reporting, and processing with R"
subtitle: "Closing remarks"
author: "Kirill Müller"
date: "2018-11-30"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
set.seed(20180516)
```

# EDA

Understand your data

- Generate questions

- Search for answers

- Rinse and repeat

---

# Important questions

- Variation

- Covariation

???

Tabular data

- Distribution
- Typica values
- Unusual values, outliers
- Missing values

---

# Discrete variables

.pull-left[
```{r outro-1, fig.height=3.5, fig.width=3.5}
ggplot(data = mpg) +
  geom_bar(
    mapping = aes(x = class)
  )
```
]

.pull-right[
```{r outro-2}
mpg %>% 
  count(class)
```
]

---

# Continuous variables

.pull-left[
```{r outro-3, fig.width=3.5, fig.height=3.5}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = displ),
    binwidth = 0.05
  )
```
]

.pull-right[
```{r outro-4, fig.width=3.5, fig.height=3.5}
ggplot(data = mpg) +
  geom_density(
    mapping = aes(x = displ)
  )
```
]

---

# Categorical vs. continuous variables

```{r outro-5, fig.height=3.5}
ggplot(data = mpg) +
  geom_density(
    mapping = aes(x = displ, y = ..scaled.., color = class),
  )
```

---

# Categorical vs. continuous variables

```{r outro-6, fig.height=3.5}
ggplot(data = mpg) +
  geom_boxplot(
    mapping = aes(x = class, y = displ),
  )
```

---

# Categorical variables

.pull-left[
```{r outro-7, fig.width = 3.5, fig.height=3.5}
ggplot(data = mpg) +
  geom_bin2d(
    mapping = aes(x = drv, y = class),
  )
```
]

.pull-right[
```{r outro-8}
mpg %>% 
  count(drv, class)
```
]

---

# Continuous variables

```{r outro-9,  fig.height=3.5}
ggplot(data = mpg) +
  geom_jitter(
    mapping = aes(x = hwy, y = cty),
    alpha = 0.3
  )
```

---

class: inverse

???

---

# More transformations

- Grouped `mutate()` and `filter()`, see r4ds 5.7.1
- Scoped functions, see `?scoped`
- `complete()` and `fill()`
- `extract()`

## Same syntax for working with databases

- *dbplyr* package
- Transformation operations are translated to SQL

---

# Joins

r4ds, chapter 13

```{r outro-10}
library(nycflights13)
flights %>%
  select(year, month, day, carrier) %>%
  left_join(airlines)
```

---

# Nested data frames

r4ds, chapter 25

```{r outro-11}
flights %>%
  nest(-month) %>% 
  arrange(month)
```

---

# Visualizations not covered

.pull-left[

## Position adjustments

jitter, dodge, stack, nudge

## Scales

labeling, color, range

]

.pull-right[

## Coordinate systems

flipping, aspect ratio, polar

## Theming

tweaking plots, standard appearance

]

---

# Extension for ggplot2

- Useful `ggplot2` extensions, https://www.ggplot2-exts.org/gallery/
    - ggstance
    - ggrepel
    - gganimate
    - GGally

---

# Pointers

Working directory hell:
- https://www.tidyverse.org/articles/2017/12/workflow-vs-script/

Symbolic link to data directory:
- Linux and OS X: `file.symlink()`
- Windows: `Sys.junction()`

R markdown: http://rmarkdown.rstudio.com/gallery.html

---

# Pointers 2

Literature:

- Quick-R [http://www.statmethods.net](http://www.statmethods.net)
- Advanced R [http://adv-r.had.co.nz](http://adv-r.had.co.nz)
- Packages suggested by RStudio [https://github.com/rstudio/RStartHere](https://github.com/rstudio/RStartHere)
