---
title: "Data visualisation, reporting, and processing with R"
subtitle: "Closing remarks"
author: "Kirill Müller"
date: "2017-10-20"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
set.seed(20180516)
```


# More transformations

- Grouped `mutate()` and `filter()`, see r4ds 5.7.1
- Scoped functions, see `?scoped`
- `complete()` and `fill()`
- `extract()`

## Same syntax for working with databases

- *dbplyr* package
- Transformation operations are translated to SQL

---

# Nested data frames

r4ds, chapter 25

```{r outro-11}
flights %>%
  nest(-month) %>% 
  arrange(month)
```

---

# Pointers

Working directory hell:
- https://www.tidyverse.org/articles/2017/12/workflow-vs-script/

Symbolic link to data directory:
- Linux and OS X: `file.symlink()`
- Windows: `Sys.junction()`

R markdown: http://rmarkdown.rstudio.com/gallery.html

---

# Pointers 2

Literature:

- Quick-R [http://www.statmethods.net](http://www.statmethods.net)
- Advanced R [http://adv-r.had.co.nz](http://adv-r.had.co.nz)
- Packages suggested by RStudio [https://github.com/rstudio/RStartHere](https://github.com/rstudio/RStartHere)
