---
title: "The pipe"
author: "Kirill Müller, cynkra GmbH"
date: "June 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)

View <- function(x) print(x)
```

## Intermediate variables

View all flights that arrived after 10:00 PM. Use an intermediate variable, a nested expression, and the pipe. Which appeals more to you?

```r
flights_after_10 <- filter(flights, ___)
View(flights_after_10)
```

```r
View(filter(flights, ___))
```

```r
flights %>%
  filter(___) %>%
  View()
```

Naming is hard!

```{r filtering-19}
early_flights <- filter(flights, dep_time >= 600)
early_late_flights <-
  filter(early_flights, arr_time >= 2200)
early_late_ua_flights <-
  filter(early_late_flights, carrier == "UA")
early_late_late_ua_flights <-
  filter(early_late_ua_flights, arr_delay > 120)
early_late_late_ua_flights_not_honolulu <-
  filter(early_late_late_ua_flights, dest != "HNL")
early_late_late_ua_flights_not_honolulu_sorted <-
  arrange(
    early_late_late_ua_flights_not_honolulu,
    distance
  )
View(early_late_late_ua_flights_not_honolulu_sorted)
```

## Nested expressions

Difficult to read.

```{r filtering-20}
View(
  arrange(
    filter(
      filter(
        filter(
          filter(
            filter(
              flights,
              dep_time <= 600
            ),
            arr_time >= 2200
          ),
          carrier == "UA"
        ),
        arr_delay > 120
      ),
      dest != "HNL"
    ),
    distance
  )
)
```

## Pipe

Extend the four solutions to view all `"UA"` flights that arrived after 10:00 PM.

```r
flights_after_10 <- filter(flights, ___)
ua_flights_after_10 <- ...
View(___)
```

```r
View(filter(filter(flights, ___)))
```

```r
flights %>%
  filter(___) %>%
  filter(___) %>%
  View()
```

```{r filtering-21}
flights %>% 
  filter(dep_time <= 600) %>% 
  filter(arr_time >= 2200) %>% 
  filter(carrier == "UA") %>% 
  filter(arr_delay > 120) %>% 
  filter(dest != "HNL") %>%
  arrange(distance) %>%
  View()
```

The original data is never updated! You still need to assign the result of a pipe to a variable:

```{r filtering-22}
late_late_ua_flights_not_honolulu <-
  flights %>% 
  filter(dep_time <= 600) %>% 
  filter(arr_time >= 2200) %>% 
  filter(carrier == "UA") %>% 
  filter(arr_delay > 120) %>% 
  filter(dest != "HNL") %>%
  arrange(distance)
```

##

Extend the four solutions to view all `"UA"` flights that departed before 6:00 AM and arrived after 10:00 PM.

##

Extend the four solutions to view all `"UA"` flights that departed before 6:00 AM and arrived after 10:00 PM and had a delay of more than two hours.

##

Extend the four solutions to view all `"UA"` flights that departed before 6:00 AM and arrived after 10:00 PM and had a delay of more than two hours, originating in one of New York City's airports.

##

Extend the four solutions to view all `"UA"` flights that departed before 6:00 AM and arrived after 10:00 PM and had a delay of more than two hours, originating in one of New York City's airports but excluding Honolulu International airport.

Hint: Consult the `airports` dataset, use a filter with the predicate `stringr::str_detect(name, "^Honolulu")` .

##

Sort the result by `distance`.
