---
title: "Interactive plots"
author: "Kirill Müller, cynkra GmbH"
date: "November 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)
```

## Interactive scatterplot

Convert one of the scatterplots you generated previously using the following pattern:

```
p <-
  data %>% 
  ___ %>%
  ggplot(aes(___)) +
    geom_point()
  
plotly::ggplotly(p)
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r plotly-1}
p <-
  mpg %>% 
  ggplot(aes(x = displ, y = hwy, color = drv)) +
    geom_jitter() +
    geom_smooth()

plotly::ggplotly(p)
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## More expressive tooltip

Experiment with the plotly controls on the top right.

Add extra aesthetics to the scatterplot. Hover with the mouse over the points:

```
p <-
  data %>% 
  ___ %>%
  ggplot(aes(___, extra1 = manufacturer, extra2 = model)) +
    geom_point()
  
plotly::ggplotly(p)
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r plotly-2}
p <-
  mpg %>% 
  ggplot(aes(x = displ, y = hwy, extra1 = manufacturer, extra2 = model)) +
    geom_jitter()

plotly::ggplotly(p)
```
</details>

## Barplot

Convert a bar plot or a histogram to plotly. How does this affect the tooltip?

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r plotly-3}
p <-
  mpg %>% 
  ggplot(aes(x = drv)) +
    geom_bar()

plotly::ggplotly(p)
```
<<<<<<< HEAD
</details>

## Crosstalk
=======
##
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

See https://plotly-book.cpsievert.me/linking-views-without-shiny.html for examples of linking two plots, or linking plots with other widgets. This requires the _crosstalk_ package:

```
install.packages("crosstalk")
```
