---
title: "Position adjustments"
author: "Kirill Müller, cynkra GmbH"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Gears

Transmission type and number of gears are encoded in the same variable, a simple bar plot helps:

```{r position-1, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = trans))
```

Visual inspection reveals that it's four gears for automatic, and five gears for manual transmission.

## Discriminate by car class

```{r position-2, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill = trans))
```

SUVs have automatic mostly, 2-seaters favor six-gear manual, compact and subcompact mostly have five-gear manual transmission. This is much easier to see with `position = "dodge"`, the different width of the bars indicates the number of different subgroups for each car class:

```{r position-3}
ggplot(data = mpg) +
  geom_bar(
    aes(x = class, fill = trans), 
    position = "dodge"
  )
```

## Boxplot

I'm using liters per 100 km as measure for fuel economy here.

```{r position-4}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = drv, y = 235 / hwy))
```

Forward drivetrains seem much more economical. Does the number of cylinders play a role? I'll try the "fill" aesthetic:

```{r position-5}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = drv, y = 235 / hwy, fill = cyl))
```

No dice. Do I also need `group =` ?

```{r position-6}
ggplot(data = mpg) +
  geom_boxplot(
    mapping = aes(
      x = drv,
      y = 235 / hwy,
      fill = cyl,
      group = cyl
    )
  )
```

The legend reveals that `cyl` is a continuous variable. I'll use its categorical equivalent, because the range is very limited.

```{r position-7}
ggplot(data = mpg) +
  geom_boxplot(
    mapping = aes(
      x = drv,
      y = 235 / hwy,
      fill = factor(cyl)
    )
  )
```

The default `position` setting looks good, let's try `"identity"`:

```{r position-8}
ggplot(data = mpg) +
  geom_boxplot(
    mapping = aes(
      x = drv,
      y = 235 / hwy,
      fill = factor(cyl)
    ),
    position = "identity"
  )
```

Doesn't look useful.
