---
title: "Position adjustments"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Gears

What's the most prevalent number of gears for manual or automatic transmissions?

```
ggplot(data = mpg, aes(x = trans)) +
  geom_bar()
```

<details>
<summary>&#x25BA; Solution:</summary>
Transmission type and number of gears are encoded in the same variable, a simple bar plot helps:

```{r position-1, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = trans))
```

Visual inspection reveals that it's four gears for automatic, and five gears for manual transmission.
</details>

## Discriminate by car class

Use the `fill` aesthetic to further discriminate by car class.
Can you explain why the parts representing transmission types are stacked on top of each other?

```
ggplot(data = mpg, mapping = aes(x = class, _____))
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r position-2, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill = trans))
```

The default value for the `position` argument to `geom_bar()` is `"stack"`.
This means that related geoms are stacked on top of each other.
</details>


## Show drivetrain types side by side

Apply a position adjustment to make it easier to detect missing combinations of drivetrain and car class.

```
ggplot(data = mpg, mapping = aes(x = class, _____)) +
  geom_bar(position = "___")
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r position-3}
ggplot(data = mpg) +
  geom_bar(
    aes(x = class, fill = trans), 
    position = "dodge"
  )
```

To use uniform width, specify `position_dodge(preserve = "single")`:

```{r position-3-bis}
ggplot(data = mpg) +
  geom_bar(
    aes(x = class, fill = trans), 
    position = position_dodge(preserve = "single")
  )
```
</details>

## Boxplot

Draw a boxplot of highway fuel economy versus drivetrain. Is fuel economy also affected by the number of cylinders?

Hint: Use `factor()` as necessary.

<details>
<summary>&#x25BA; Solution:</summary>
I'm using liters per 100 km as measure for fuel economy here.

```{r position-4}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = drv, y = 235 / hwy))
```

Forward drivetrains seem much more economical. Does the number of cylinders play a role? I'll try the "fill" aesthetic:

```{r position-5}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = drv, y = 235 / hwy, fill = cyl))
```

No dice. Do I also need `group =` ?

```{r position-6}
ggplot(data = mpg) +
  geom_boxplot(
    mapping = aes(
      x = drv,
      y = 235 / hwy,
      fill = cyl,
      group = cyl
    )
  )
```

The legend reveals that `cyl` is a continuous variable. I'll use its categorical equivalent, because the range is very limited.

```{r position-7}
ggplot(data = mpg) +
  geom_boxplot(
    mapping = aes(
      x = drv,
      y = 235 / hwy,
      fill = factor(cyl)
    )
  )
```

The default `position` setting looks good, we use `position_dodge(preserve = "single")` again for uniform width:

```{r position-8}
ggplot(data = mpg) +
  geom_boxplot(
    mapping = aes(
      x = drv,
      y = 235 / hwy,
      fill = factor(cyl)
    ),
    position = position_dodge(preserve = "single")
  )
```

Doesn't look useful.
</details>

## More exercises

Find more exercises in Section 3.8.1 of r4ds.
