---
title: "R + markdown"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r rmarkdown-start-1, echo = FALSE, results = "asis"}
text <- '``````\n---
title: "Test document"
author: "Kirill Müller, cynkra GmbH"
output: html_document\n---
\n```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)\n```

# Introduction

We will be reviewing a dataset of `r nrow(mpg)` cars. The dataset is listed below.
\n```{r}
mpg\n```

## Relationship between engine displacement and highway mileage

The plot below should *not* surprise us: The larger the engine, the more energy it consumes per distance driven (in $\\frac{\\text{J}}{\\text{km}}$), the less is the mileage per unit of gasoline (in $\\frac{\\text{miles}}{\\text{gallon}}$).
\n```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()\n```

Possible reasons why people would choose a larger engine include:

- ...
- ...
- ...\n``````'

knitr::asis_output(text)
```


# Introduction

We will be reviewing a dataset of `r nrow(mpg)` cars. The dataset is listed below.

```{r rmarkdown-start-2}
mpg
```

## Relationship between engine displacement and highway mileage

The plot below should *not* surprise us: The larger the engine, the more energy it consumes per distance driven (in $\frac{\text{J}}{\text{km}}$), the less is the mileage per unit of gasoline (in $\frac{\text{miles}}{\text{gallon}}$).

```{r rmarkdown-start-3}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

Possible reasons why people would choose a larger engine include:

- ...
- ...
- ...
