---
title: "Scales"
author: "Kirill Müller, cynkra GmbH"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(20180516)
knitr::opts_chunk$set(cache = TRUE)
```

## Naming scales

Choose one or two of your previous plots, and give the scales a proper name.

Hint: Use one or more `scale_...()` functions.
   
```
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = factor(year))
) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~class) +
  scale_x_continuous(name = "___") +
  scale_y_continuous(name = "___") +
  scale_color_discrete(name = "___")
```

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
Reusing a plot from the previous exercise:

```{r scales-1, echo=TRUE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = factor(year))
  ) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = factor(year)),
    method = "lm"
  ) +
  facet_wrap(~class) +
  scale_x_continuous(name = "Displacement") +
  scale_y_continuous(name = "Fuel economy [miles per gallon]") +
  scale_color_discrete(name = "Year")
```

There exist shortcuts `xlab()` and `ylab()` for x and y labels, but not for color or fill. I recommend to stick with the explicit versions.
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## Changing scales twice

What happens if you call the same `scale_...()` function twice in a plot?

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
A warning occurs:

```{r scales-2, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  scale_x_continuous(name = "Displacement") +
  scale_x_continuous(name = "Displacement")
```
<<<<<<< HEAD
</details>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

## ColorBrewer scales

Add an overview over all available colorblind friendly Brewer scales to your rmarkdown document.

Hint: Insert a code chunk containing `RColorBrewer::display.brewer.all(colorblindFriendly = TRUE)`

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r scales-3}
RColorBrewer::display.brewer.all(colorblindFriendly = TRUE)
```
</details>

## A better scale for car class

Choose an appealing scale for one of your plots that use the "color" or "fill" aesthetics. Apply it with `scale_..._brewer("<name>", palette = "<palette>")`

Hint: Use `scale_..._distiller()` if you have mapped a continuous variable.

<<<<<<< HEAD
<details>
<summary>&#x25BA; Solution:</summary>
=======
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815
```{r scales-4, echo=TRUE}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = displ, y = hwy, color = class)) +
  scale_x_continuous(name = "Displacement") +
  scale_y_continuous(name = "Fuel economy [miles per gallon]") +
  scale_color_brewer(name = "Car class", palette = "Set2")
```
<<<<<<< HEAD
</details>

## Practice
=======

##
>>>>>>> 0ce8a33a43c4959d6f16c508b69e30907a3dd815

Practice Markdown: Create sections, subsections, item lists, emphasized and bold text, ... .
