---
title: "Scatterplots"
author: "Kirill Müller, Patrick Schratz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Show mpg on console

Print the `mpg` dataset to your console.

<details>
<summary>&#x25BA; Solution:</summary>
```{r scatter-1}
mpg
```
</details>

## RStudio's view pane

View `mpg` in RStudio's dataset pane (via `View()`). Display all Audis sorted by year then cylinder. What is the maximum number of highway miles per gallon in this dataset?

```
view(___)
```

<details>
<summary>&#x25BA; Solution:</summary>
```r
view(mpg)
```

- Filtering works by clicking on the "Filter" button, then entering a text (for text columns) or restricting the range of values (for numeric columns).
- Arranging works as usual by clicking on a column header. Because the sort operation is stable, you can sort by two variables by first sorting by the *second* variable, then by the *first*.
- The 1999 VW Jetta 1.9 and VW New Beetle 1.9 have the best highway economy, both will drive you 44 miles for each gallon of fuel.
- Note that the view pane currently shows at most 100 columns.
</details>

## Scatterplots

Create a few scatterplots using the following template (here with `hwy` versus `displ`):

```{r echo=TRUE, eval=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
    
Share particularly interesting or boring examples you might encounter.  Did you find anything surprising?

<details>
<summary>&#x25BA; Solution:</summary>
```{r scatter-2, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

More examples:

```{r scatter-3}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = drv))
```

```{r scatter-4}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = year, y = cyl))
```

```{r scatter-5}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = manufacturer))
```

```{r scatter-6}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy))
```
</details>

## Convert units

Can you plot highway fuel economy given as liters per 100 kilometers against engine displacement?

Hint: Use the formula `235 / hwy` to convert from miles per gallon.

```
ggplot(data = ___) +
  geom_point(mapping = aes(x = displ, y = ___))
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r scatter-7, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = 235 / hwy))
```
</details>

## More exercises

Find more examples in Section 3.2.4 of the "R for data science" book (http://r4ds.had.co.nz).
