---
title: "Scatterplots with three variables"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Map to the color aesthetic

In the `hwy` vs. `displ` plot, map an additional variable to the "color" aesthetic.
Which cars consume more fuel than expected by the general trend?

```
ggplot(_____) +
  geom_point(mapping = aes(x = ___, y = ___, color = ___))
```

<details>
<summary>&#x25BA; Solution:</summary>
Simply insert `color = <var>` in the `aes()` call, where `<var>` is a variable in the `mpg` dataset:

```{r scatter3-1, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

```{r scatter3-2, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl))
```

```{r scatter3-3, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = trans))
```

```{r scatter3-4, echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))
```
</details>

## Map to other aesthetics

Experiment with the "color", "shape", "size", and "alpha" aesthetics.  Which combinations of attribute class (categorical/continuous) and aesthetics work well, which don't? Expand on the more surprising examples in the previous exercises.

Hint: Use `factor(year)` to convert continuous variables with a limited set of values to categorical variables.

```
ggplot(_____) +
  geom_point(mapping = aes(x = ___, y = ___, ___ = ___))
```

<details>
<summary>&#x25BA; Solution:</summary>
Looks like the cars with large displacement are two-seaters with rear drivetrain.

The following won't work, because `cyl` is stored as a continuous variable:

```{r scatter3-5-bad, error = TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = cyl))
```

The `factor()` function helps:

```{r scatter3-5}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = factor(cyl)))
```

On the other hand, a continuous variable is fine for the "size" aesthetic...

```{r scatter3-6}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = cyl))
```

...which isn't recommended (but still works) for a categorical variable:

```{r scatter3-7}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = factor(cyl)))
```

The "alpha" aesthetic controls transparency, and accepts both kinds of variable:

```{r scatter3-8}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = cyl))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = factor(cyl)))
```
</details>

## Change more than one aesthetic

Can you change both color and shape at the same time? What about the other aesthetics?

```
ggplot(_____) +
  geom_point(mapping = aes(x = ___, y = ___, ___ = ___, ___ = ___))
```

<details>
<summary>&#x25BA; Solution:</summary>
You can change any number of properties of a geom at your discretion, this is one of **ggplot2**'s strenghts.

```{r scatter3-9}
ggplot(data = mpg) +
  geom_point(mapping = aes(
    x = displ,
    y = hwy,
    color = class,
    size = cyl,
    shape = drv,
    alpha = fl
  ))
```

A legend is added automatically.
It may not be as useful to change all properties at the same time, but it's very easy to assign a variable to a different aesthetic, or use another variable for that aesthetic.
</details>

## Map a variable to more than one aesthetic

What happens if you map the same variable to more than one aesthetic?

```
ggplot(_____) +
  geom_point(mapping = aes(x = ___, y = ___, ___ = ___, ___ = ___))
```

<details>
<summary>&#x25BA; Solution:</summary>
There are no restrictions on how many times a variable is mapped in any given plot:

```{r scatter3-10}
ggplot(data = mpg) +
  geom_point(mapping = aes(
    x = displ,
    y = hwy,
    color = displ,
    size = displ,
    alpha = displ
  ))
```

Mapping the same variable to two aesthetics is sometimes useful to further emphasize this variable.
</details>

## More exercises

Find more exercises in Section 3.3.1 of r4ds.
