---
title: "Scatterplots with three variables"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Map to the color aesthetic

Simply insert `color = <var>` in the `aes()` call, where `<var>` is a variable in the `mpg` dataset:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl))
```

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = trans))
```

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))
```

Looks like the cars with large displacement are two-seaters with rear drivetrain.


## Map to other aesthetics

The following won't work, because `cyl` is stored as a continuous variable:

```r
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = cyl))
```

The `factor()` function helps:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = factor(cyl)))
```

On the other hand, a continuous variable is fine for the "size" aesthetic...

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = cyl))
```

...which isn't recommended (but still works) for a categorical variable:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = factor(cyl)))
```

The "alpha" aesthetic controls transparency, and accepts both kinds of variable:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = cyl))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = factor(cyl)))
```


## Change more than one aesthetic

You can change any number of properties of a geom at your discretion, this is one of **ggplot2**'s strenghts.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(
    x = displ,
    y = hwy,
    color = class,
    size = cyl,
    shape = drv,
    alpha = fl
  ))
```

A legend is added automatically.  It may not be as useful to change all properties at the same time, but it's very easy to assign a variable to a different aesthetic, or use another variable for that aesthetic.


## Map a variable to more than one aesthetic

There are no restrictions on how many times a variable is mapped in any given plot:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(
    x = displ,
    y = hwy,
    color = displ,
    size = displ,
    alpha = displ
  ))
```

Mapping the same variable to two aesthetics is sometimes useful to further emphasize this variable.
