---
title: "Select and rename"
author: "Kirill Müller, cynkra GmbH"
date: "June 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)
```

## Select first five variables

Find three ways to select the first five variables from the `flights` dataset.

```r
flights %>% 
  select(___, ___, ________)

flights %>% 
  select(___:___)

flights %>% 
  select(___:___)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r select-rename-1}
flights %>%
  select(year, month, day, dep_time, sched_dep_time)
```

```{r select-rename-2}
flights %>%
  select(year:sched_dep_time)
```

```{r select-rename-3}
## Numeric indexes work, too

flights %>%
  select(1:5)
```
</details>

## Exclude the date

Find three ways to exclude the date of the flight.

```r
flights %>% 
  select(___, ___, ______________________)

flights %>% 
  select(-___, -___, -___)

flights %>% 
  select(-___:-___)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r select-rename-4}
flights %>%
  select(-year, -month, -day)
```

```{r select-rename-5}
flights %>%
  select(-year:-day)
```

```{r select-rename-6}
## Numeric indexes work, too

flights %>%
  select(-1:-3)
```
</details>

## Departure variables

Select all variables related to departure.

```r
flights %>% 
  select(___, ___, _______)

flights %>% 
  select(starts_with("___"))
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r select-rename-7}
## contains()

flights %>% 
  select(contains("dep_"))
```
</details>

## Move departure variables to end

Move the variables related to scheduled time to the end of the table.

```r
flights %>% 
  select(-___, -___, _______, everything(), ___, ___)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r select-rename-8}
## everything()

flights %>%
  select(-contains("dep_"), everything(), contains("dep_"))
```
</details>

## Contour plot

Create a contour plot of departure and arrival time. Rename the columns to show prettily in the plot. Restrict the plot to all flights that arrive before 5:00 AM.

```r
flights %>%
  select(`___` = ___, `___` = ___, ___) %>%
  filter(___) %>%
  ggplot(aes(x = `___`, y = `___`)) +
  geom_density2d()
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r select-rename-9}
## Overriding label names

flights %>% 
  ggplot() +
  geom_density2d(aes(dep_time, arr_time)) +
  scale_x_continuous(name = "Departure time") +
  scale_y_continuous(name = "Arrival time")
```

```{r select-rename-10}
## Renaming in the data

flights %>% 
  rename(`Departure time` = dep_time) %>%
  rename(`Arrival time` = arr_time) %>%
  ggplot() +
  geom_density2d(aes(`Departure time`, `Arrival time`))
```

```{r select-rename-11}
flights %>% 
  filter(arr_time < 500) %>%
  rename(`Departure time` = dep_time) %>%
  rename(`Arrival time` = arr_time) %>%
  ggplot() +
  geom_density2d(aes(`Departure time`, `Arrival time`))
```

```{r select-rename-12}
## Use coord_fixed() for fixing axes
flights %>% 
  filter(arr_time < 500) %>%
  rename(`Departure time` = dep_time) %>%
  rename(`Arrival time` = arr_time) %>%
  ggplot() +
  geom_density2d(aes(`Departure time`, `Arrival time`)) +
  coord_fixed()
```
</details>

## More exercises

Find more exercises in Section 5.4.1 of r4ds.
