---
title: "Spreading and gathering"
author: "Kirill Müller, cynkra GmbH"
date: "June 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(20180516)
knitr::opts_chunk$set(cache = TRUE)
```

## table2 to table1

Use `spread()` to convert `table2` to `table1`. What is the meaning of the `key` and `value` arguments?

```r
table2 %>%
  spread(_____)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r spread-gather-1}
table2 %>% 
  spread(type, count)
```
</details>

## table1 to table2

Use `gather()` to convert `table1` to `table2`. Try an inclusive and an exclusive selection. Do you need an extra transformation to make the result fully identical? Can you reuse `key` and `value` from the previous result?

```r
table1 %>%
  gather(_____, ___:___)

table1 %>%
  gather(_____, -___:-___)
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r spread-gather-2}
table1 %>% 
  gather(type, count, cases, population)
```

```{r spread-gather-3}
table2 %>% 
  gather(type, count, -country, -year)
```

Are the two calls symmetrical?

No, we need to arrange this result:

```{r spread-gather-4}
table1 %>% 
  gather(type, count, -country, -year) %>% 
  arrange(country, year, type)
```
</details>

## Plot table-x

Visualize the data: plot cases, population, and both. Which of `table1` or `table2` is more suitable in which case?

```r
___ %>%
  ggplot(aes(___)) +
  geom_col()

___ %>%
  ggplot(aes(___)) +
  geom_col() +
  facet_grid(___ ~ ___, scales = "free")

___ %>%
  ggplot(aes(___, ___)) +
  geom_point()
```

<details>
<summary>&#x25BA; Solution:</summary>
For showing one measurement:

```{r spread-gather-5}
table1 %>%
  ggplot(aes(country, population, fill = factor(year))) +
  geom_col(position = "dodge")
```

For mapping measurement type to an aesthetic:

```{r spread-gather-6}
table2 %>%
  ggplot(aes(country, count, fill = type)) +
  geom_col() +
  facet_wrap(~year)
```

For mapping both measurements to two aesthetics:

```{r spread-gather-7}
table1 %>%
  ggplot(aes(population, cases, color = country)) +
  geom_point() +
  facet_wrap(~year)
```

Can also use `table1` to show only one measurement:

```{r spread-gather-8}
table2 %>%
  filter(type == "cases") %>%
  ggplot() +
  geom_col(aes(country, count, fill = type)) +
  facet_wrap(~year)
```
</details>

## Binding

Use `gather()` to convert `table4a` and `table4b` to `table2`. Can you do the same with just one `gather()` call?

Hint: Use `bind_rows()` to combine similar tibbles.

```r
cases_tbl <-
  table4a %>%
  gather(_____) %>%
  mutate(type = "cases")

population_tbl <-
  table4a %>%
  gather(_____) %>%
  mutate(___)
  
bind_rows(_____) %>%
  _____ %>%
  _____
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r spread-gather-9}
cases_tbl <-
  table4a %>%
  gather(year, count, -country) %>%
  mutate(type = "cases")

population_tbl <-
  table4b %>% 
  gather(year, count, -country) %>%
  mutate(type = "population")

bind_rows(cases_tbl, population_tbl) %>%
  select(country, year, everything()) %>% 
  arrange(country, year, type)
```

With just one call to `gather()`:

```{r spread-gather-10}
cases_tbl <-
  table4a %>%
  mutate(type = "cases")

population_tbl <-
  table4b %>% 
  mutate(type = "population")

bind_rows(cases_tbl, population_tbl) %>%
  gather(year, count, -country, -type) %>%
  select(country, year, everything()) %>% 
  arrange(country, year, type)
```

Shorter, with the `.id` argument to `bind_rows()`:

```{r spread-gather-10-a}
bind_rows(cases = table4a, population = table4b, .id = "type") %>%
  gather(year, count, -country, -type) %>%
  select(country, year, everything()) %>% 
  arrange(country, year, type)
```
</details>

## Economies

Create a scatterplot from the `mpg` dataset that shows both highway and city fuel economy against engine displacement with two different colors using only one `geom_point()` call.

```r
mpg %>%
  _____ %>%
  ggplot(aes(x = displ, y = ___)) +
  geom_point()
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r spread-gather-11}
mpg %>%
  select(displ, hwy, cty) %>%
  gather(cty_hwy, value, hwy, cty) %>%
  ggplot() +
  geom_jitter(aes(displ, value, color = cty_hwy))
```
</details>

## More exercises

Find more exercises in Section 12.3.3 of r4ds.
