---
title: "Statistical transformations"
author: "Kirill Müller"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(cache = TRUE)
```

## Cars by drivetrain

I tried `geom_histogram()` and `geom_col()`, neither worked. The histogram is for continuous data only, for `geom_col()` I'd need to supply actual counts which I don't have. The `geom_bar()` function computes the counts for me by applying the `"count"` statistical transformation to my data before plotting.

We need only the "x" aesthetic, "y" is computed automatically. `drv` is the relevant variable.

```{r}
ggplot(mpg) +
  geom_bar(aes(x = drv))
```

## Explicit count

The output is the same:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..count..))
```

By default, `geom_bar()` plots the computed "count" aesthetic, which is by convention written with two surrounding dots.


## Proportions

Plotting overall proportions changes the y scale:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..prop.., group = 1))
```

Omitting `group` will plot proportions per drivetrain in each column, which are 1 by default:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, y = ..prop..))
```


## Fuel economy

We use a histogram of `hwy`, again only the "x" aesthetic needs to be supplied:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = hwy))
```

The bin width should be set explicitly to avoid the warning:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy),
    binwidth = 1
  )
```

The frequency polygon is very similar. We tweak aesthetics manually in order to be able to distinguish between both layers.

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy),
    fill = NA,
    color = "blue",
    binwidth = 1
  ) +
  geom_freqpoly(
    mapping = aes(x = hwy),
    binwidth = 1,
    size = 2
  )
```


## Fuel economy by drivetrain

We can set the "fill" aesthetic:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy, fill = drv),
    binwidth = 1
  )
```

However, some bars are hidden. This happens less if we use color instead:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_histogram(
    mapping = aes(x = hwy, color = drv),
    fill = NA,
    binwidth = 1
  )
```

The frequency polygon offers a much clearer display of the same data:

```{r echo=TRUE}
ggplot(data = mpg) +
  geom_freqpoly(
    mapping = aes(x = hwy, color = drv),
    binwidth = 1
  )
```

## Fuel economy per drivetrain (a history of failed attempts)

Can I just copy and adapt the code from the last solution?

```{r}
ggplot(mpg, aes(hwy, color = drv)) +
  geom_histogram(fill = NA, color = "black", size = 0.2, binwidth = 2) +
  geom_freqpoly(color = "red", binwidth = 2)
```

Nope. Something's wrong here. Maybe if I follow the template?

```{r}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv),
    binwidth = 2,
    fill = NA,
    color = "black",
    size = 0.2
  )
```

I can't use color both as variable and manual aesthetic, ggplot2 silently uses the manual assignment! Need to be careful with assigning aesthetics.


```{r}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv, group = drv),
    binwidth = 2,
    size = 0.2,
    fill = NA
  ) +
  geom_freqpoly(
    aes(hwy, color = drv, group = drv),
    binwidth = 2
  )
```



```{r}
ggplot(mpg) +
  geom_histogram(
    aes(hwy, color = drv),
    binwidth = 2,
    size = 0.2,
    fill = NA
  )
```


```{r}
ggplot(mpg) +
  geom_freqpoly(
    aes(hwy, color = drv),
    binwidth = 2
  )
```

