---
title: "Summarize by multiple variables"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)
```

## Distinct airlines per relation

Which relation is serviced by the largest number of distinct airlines? Find a solution using `summarize()`, one using `count()`, and one using `tally()`. Which is more elegant?

```r
flights %>%
  group_by(___, ___, airline) %>%
  summarize(n = n()) %>%
  summarize(n_airlines = ___) %>%
  ungroup() %>%
  arrange(___) %>%
  head(1)

flights %>%
  count(_____) %>%
  count(_____) %>%
  _____ %>%
  _____

flights %>%
  group_by(_____) %>%
  tally() %>%
  tally(wt = NULL) %>%
  _____ %>%
  _____
```

```{r summarize-mult-1}
flights %>%
  group_by(origin, dest, carrier) %>%
  summarize(n_flights = n()) %>%
  summarize(n_distinct_carriers = n()) %>%
  ungroup() %>% 
  arrange(desc(n_distinct_carriers))
```

Much shorter:

```{r summarize-mult-2}
flights %>%
  count(origin, dest, carrier) %>%
  count(origin, dest) %>% 
  ungroup() %>% 
  arrange(desc(nn))
```

Alternatively:

```{r summarize-mult-3}
flights %>%
  group_by(origin, dest, carrier) %>%
  tally() %>%
  tally(wt = NULL) %>%
  ungroup() %>%
  arrange(desc(nn))
```

## Cancelled flights per month per airline

Compute the share of cancelled flights per month per airline.

```r
flights %>% 
  group_by(_____) %>% 
  summarize(share_of_cancelled = _____) %>%
  ungroup()
```

```{r summarize-mult-4}
cancelled_flights <-
  flights %>% 
  group_by(carrier, month) %>% 
  summarize(share_of_cancelled = mean(is.na(dep_time))) %>%
  ungroup()

cancelled_flights
```

## Heat map of cancelled flights

Create a heat map of cancelled flights.

```r
cancelled_flights <-
  _____

cancelled_flights %>% 
  ggplot() +
  geom_raster(
    aes(
      x = ___,
      y = factor(month),
      fill = ___
    )
  )
```

```{r summarize-mult-5}
cancelled_flights %>%
  ggplot() +
  geom_raster(aes(x = carrier, y = factor(month), fill = share_of_cancelled))
```

##

Find more exercises in Section 5.6.7 of r4ds.
