---
title: "Summarize"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)
```

## Useful summary functions

- `sum()`, `prod()`
- `mean()`, `median()`
- `sd()`, `IQR()`, `mad()`
- `min()`, `quantile()`, `max()`
- `n()`
- `sum()` and `mean()` for logical variables
- `first()`, `last()`, `nth()`

Don't forget `na.rm = TRUE` if needed!

## Mean arrival and dep delay

Compute the mean arrival and departure delay overall, and per origin airport. What is the standard deviation of these variables? What is New York City's busiest airport?

```r
flights %>%
  summarize(mean(___, na.rm = ___))

flights %>%
  group_by(___) %>%
  summarize(___)

flights %>%
  count(___) %>%
  arrange(___)
```

```{r summarize-1}
flights %>%
  summarize(mean(arr_delay, na.rm = TRUE))
```

```{r summarize-2}
flights %>%
  summarize(mean(dep_delay, na.rm = TRUE))
```

```{r summarize-3}
flights %>%
  summarize(
    mean(arr_delay, na.rm = TRUE),
    mean(dep_delay, na.rm = TRUE),
    sd(arr_delay, na.rm = TRUE),
    sd(dep_delay, na.rm = TRUE)
  )
```

```{r summarize-4}
flights %>%
  summarize_at(
    vars(arr_delay, dep_delay),
    funs(mean(., na.rm = TRUE), sd(., na.rm = TRUE))
  )
```

```{r summarize-5}
flights %>%
  group_by(origin) %>%
  summarize(
    mean(arr_delay, na.rm = TRUE),
    mean(dep_delay, na.rm = TRUE)
  )
```

```{r summarize-6}
flights %>%
  count(origin) %>%
  arrange(desc(n))
```

```{r summarize-6-1}
flights %>%
  count(origin, sort = TRUE)
```

## Air time by carrier

Which carriers had the longest accumulated air time, excluding cancelled flights?

```r
flights %>%
  group_by(___) %>%
  summarize(acc_air_time = sum(_____)) %>% 
  ungroup()
```

```{r summarize-7}
total_airtime_by_carrier <-
  flights %>% 
    group_by(carrier) %>% 
    summarize(acc_air_time = sum(air_time, na.rm = TRUE)) %>% 
    ungroup()

total_airtime_by_carrier
```

Plot a bar chart of the accumulated air time per airline, with a suitable unit for the total time.

Hint: Use `forcats::fct_inorder()` to fix the ordering of a categorical variable before plotting.

```r
total_airtime_by_carrier <-
  _____
  
total_airtime_by_carrier %>%
  arrange(acc_air_time) %>%
  mutate(carrier = forcats::fct_inorder(carrier)) %>%
  ggplot(aes(___)) +
    geom_col()
```

```{r summarize-8}
total_airtime_by_carrier %>% 
  arrange(acc_air_time) %>%
  mutate(carrier = forcats::fct_inorder(carrier)) %>%
  ggplot() +
  geom_col(aes(carrier, acc_air_time / 60 / 24 / 365)) +
  coord_flip()
```

## Carriers for long-distance routes

Which carriers specialize on long-distance routes?

```r
flights
  _____ %>%
  _____ %>%
  _____ %>%
  arrange(___)
```

```{r summarize-9}
carrier_arranged_by_mean_distance <-
  flights %>%
  group_by(carrier) %>% 
  summarize(mean_distance = mean(distance)) %>% 
  ungroup() %>%
  arrange(desc(mean_distance))

carrier_arranged_by_mean_distance
```

Plot a bar chart of median distance per flight. Do you use `geom_bar()` or `geom_col()`? Why?

```r
median_miles_by_carrier <-
  _____ %>%
  _____ %>%
  _____

median_miles_by_carrier %>%
  arrange(_____) %>%
  mutate(_____) %>%
  ggplot(aes(___)) +
    geom____()
```

```{r summarize-10}
carrier_arranged_by_mean_distance %>% 
  ggplot() +
  geom_col(aes(x = forcats::fct_inorder(carrier), y = mean_distance))
```

## Worst plane

Which plane had the most failed departure attempts? Can you find a solution without `filter()`?

Hint: Use the idiom `sum(___)` to count the rows where a predicate is true.

```r
flights %>%
  filter(is.na(dep_time)) %>%
  group_by(tailnum) %>%
  _____ %>%
  _____ %>%
  filter(!is.na(tailnum)) %>%
  arrange(desc(___)) %>% 
  head(1)

# Alternative without filter():
flights %>%
  group_by(tailnum) %>%
  _____ %>%
  _____ %>%
  arrange(_____) %>%
  head(1)
```

```{r summarize-11}
flights %>% 
  filter(is.na(dep_time)) %>%
  group_by(tailnum) %>% 
  summarize(not_departed = n()) %>% 
  ungroup() %>% 
  filter(!is.na(tailnum)) %>%
  arrange(desc(not_departed))
```

Compute the ratio of short-distance routes (less than 300 miles) for each airline.

Hint: Use the idiom `mean(___)` to compute the share of rows where a predicate is true.

```r
flights %>%
  group_by(carrier) %>%
  _____ %>% 
  ungroup()
```

```{r summarize-12}
flights %>% 
  group_by(tailnum) %>% 
  summarize(not_departed = sum(is.na(dep_time))) %>% 
  ungroup()
```

```{r summarize-13}
flights %>% 
  filter(is.na(dep_time)) %>%
  count(tailnum, sort = TRUE)
```

## Short-distance routes per airline

Plot a bar chart of the ratio of short-distance routes.

```r
short_distance_route_ratio <-
  _____

short_distance_route_ratio %>%
  ggplot(aes(___)) +
    geom_col()
```

```{r summarize-14}
short_distance_route_ratio <-
  flights %>%
  group_by(carrier) %>% 
  summarize(short_distance_ratio = mean(distance < 300)) %>% 
  ungroup() %>%
  arrange(desc(short_distance_ratio))

short_distance_route_ratio

short_distance_route_ratio %>%
  ggplot() +
  geom_col(aes(x = forcats::fct_inorder(carrier), y = short_distance_ratio))
```

##

Find more exercises in item 1 of Section 5.6.7 of r4ds.
