---
title: "Theming"
author: "Kirill Müller, cynkra GmbH"
date: "June 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(20180516)
knitr::opts_chunk$set(cache = TRUE)
```

## Favorite theme

Choose your favorite among the predefined themes.

Hint: All start with `theme_...()`, but watch out for `theme_set()`.

For presentations, `theme_bw()` or `theme_minimal()` look useful:

```{r theme-1, echo=TRUE}
ggplot(data = mpg) +
  theme_bw() +
  geom_jitter(mapping = aes(x = displ, y = hwy, color = class)) +
  scale_x_continuous(name = "Displacement") +
  scale_y_continuous(name = "Fuel economy [miles per gallon]") +
  scale_color_brewer(name = "Car class", palette = "Set2")

ggplot(data = mpg) +
  theme_minimal() +
  geom_jitter(mapping = aes(x = displ, y = hwy, color = class)) +
  scale_x_continuous(name = "Displacement") +
  scale_y_continuous(name = "Fuel economy [miles per gallon]") +
  scale_color_brewer(name = "Car class", palette = "Set2")
```

## Legend at bottom

Apply `theme(legend.position = "bottom")` to a plot with a color or fill legend.  What happens if you then apply `theme_bw()`? Why?

```{r theme-2, echo=TRUE}
ggplot(data = mpg) +
  theme_bw() +
  geom_jitter(mapping = aes(x = displ, y = hwy, color = class)) +
  scale_x_continuous(name = "Displacement") +
  scale_y_continuous(name = "Fuel economy [miles per gallon]") +
  scale_color_brewer(name = "Car class", palette = "Set2") +
  theme(legend.position = "bottom")
```

`theme_bw()` overwrites everything, including the legend position:

```{r theme-3, echo=TRUE}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = displ, y = hwy, color = class)) +
  scale_x_continuous(name = "Displacement") +
  scale_y_continuous(name = "Fuel economy [miles per gallon]") +
  scale_color_brewer(name = "Car class", palette = "Set2") +
  theme(legend.position = "bottom") +
  theme_bw()
```
