---
title: "Window functions"
author: "Kirill Müller, cynkra GmbH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(cache = TRUE)
```

## Busiest month

Which month is busiest in terms of miles flown, per carrier?

Hint: Compute the share of yearly miles flown of each airline in each month.

```r
flights %>%
  group_by(___, ___) %>%
  summarize(total_distance_by_carrier = sum(distance)) %>%
  mutate(total_distance = sum(___)) %>%
  ungroup() %>%
  mutate(month_share_by_carrier = ___ / ___) %>% 
  arrange(month_share_by_carrier) %>% 
  group_by(___) %>%
  slice(1) %>%
  ungroup()
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r window-1}
monthly_shares <-
  flights %>%
  group_by(carrier, month) %>%
  summarize(distance = sum(distance)) %>%
  mutate(total_distance = sum(distance)) %>%
  ungroup() %>%
  mutate(month_share = distance / total_distance) %>%
  arrange(month_share_by_carrier) %>% 
  group_by(carrier) %>%
  slice(1) %>%
  ungroup()

monthly_shares
```
</details>

## Heat map of miles flown

Draw a heat map of miles flown per month per airline to see if this pattern holds across all airlines.

```r
monthly_shares <-
  _____

monthly_shares %>%
  ggplot(aes(factor(month), ___, fill = ___)) +
  geom_tile() +
  scale_fill_continuous(trans = "log10")
```

<details>
<summary>&#x25BA; Solution:</summary>
```{r window-1-a}
monthly_shares %>% 
  ggplot(aes(factor(month), carrier, fill = month_share)) +
  geom_tile() +
  scale_fill_continuous(trans = "log10")
```
</details>

## Busiest month

Which month is busiest in terms of miles flown, over all carriers?

```r
flights %>%
  group_by(___) %>%
  mutate(total_distance = sum(___)) %>%
  mutate(month_share = ___ / ___) %>% 
  arrange(desc(month_share)) %>%
  slice(1)
```

## Busiest month, visualized

Visualize with a bar chart.

## More Examples

Find more exercises in Section 5.7.1 of r4ds.
