## Extensions

A mass of R packages extending {ggplot2} exists.
Many are listed at http://www.ggplot2-exts.org/gallery/.

Here is a selected list of our favorite {ggplot2} extensions including some use examples.

{ggsci}: https://nanx.me/ggsci/

{ggforce}: https://ggforce.data-imaginist.com/

{patchwork}: https://patchwork.data-imaginist.com/

{gganimate}: https://gganimate.com/

{ggtext}: https://github.com/clauswilke/ggtext

{ggiraph}: http://davidgohel.github.io/ggiraph

{ggbeeswarm}: https://github.com/eclarke/ggbeeswarm

{esquisse}: https://dreamrs.github.io/esquisse

( {ggstatsplot}: https://indrajeetpatil.github.io/ggstatsplot )

( {ggedit}: https://github.com/metrumresearchgroup/ggedit )

( {lindia}: https://github.com/yeukyul/lindia )

<details><summary>*Click here to show setup code.*</summary>
```{r setup-vis-adv}
library(tidyverse)
library(ggsci)
library(ggpubr)
library(patchwork)
library(ggpmisc)
library(ggiraph)
library(ggbeeswarm)
library(gganimate)
library(ggrepel)
library(ggforce)
library(gapminder)
```
</details>

<br>

### {ggsci}

```{r ggsci}
p1 <- ggplot(mpg, aes(manufacturer)) +
  geom_bar(aes(fill = fl))

p2 <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = fl))
```

```{r patchwork, fig.width=8, fig.height=4}
library("patchwork")

p1_npg <- p1 + ggsci::scale_fill_npg()
p2_nejm <- p2 + ggsci::scale_color_nejm()

p1_npg + p2_nejm
```

### {ggforce}

```{r ggforce, fig.width=8, fig.height=5}
ggplot(iris, aes(Petal.Length, Petal.Width, colour = Species)) +
  geom_point() +
  ggforce::facet_zoom(x = Species == "versicolor")
```

### {gganimate}

```{r gganimate, fig.width=8, fig.height=5, message=FALSE}
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  gganimate::transition_time(year) +
  ease_aes('linear')
```

### {ggtext}

```{r ggtext-1}
df <- data.frame(
  label = c(
    "Some text **in bold.**",
    "Linebreaks<br>Linebreaks<br>Linebreaks",
    "*x*<sup>2</sup> + 5*x* + *C*<sub>*i*</sub>",
    "Some <span style='color:blue'>blue text **in bold.**</span><br>And *italics text.*<br>
    And some <span style='font-size:18pt; color:black'>large</span> text."
  ),
  x = c(.2, .1, .5, .9),
  y = c(.8, .4, .1, .5),
  hjust = c(0.5, 0, 0, 1),
  vjust = c(0.5, 1, 0, 0.5),
  angle = c(0, 0, 45, -45),
  color = c("black", "blue", "black", "red"),
  fill = c("cornsilk", "white", "lightblue1", "white")
)
```

```{r ggtext-2, fig.width=8, fig.height=5}
ggplot(df) +
  aes(
    x, y,
    label = label, angle = angle, color = color,
    hjust = hjust, vjust = vjust
  ) +
  ggtext::geom_richtext(
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding
  ) +
  geom_point(color = "black", size = 2) +
  scale_color_nejm() +
  xlim(0, 1) + ylim(0, 1) +
  theme_pubr()
```

### {ggrepel}

```{r ggrepel-1}
no_repel <- ggplot(mtcars, aes(wt, mpg)) +
  geom_text(label = rownames(mtcars), size = 3) +
  geom_point(color = "red") +
  theme_pubr()
```


```{r ggrepel-2}
with_repel <- ggplot(mtcars, aes(wt, mpg)) +
  ggrepel::geom_text_repel(label = rownames(mtcars), size = 3) +
  geom_point(color = "red") +
  theme_pubr()
```

```{r, fig.width=9, fig.height=5}
no_repel + with_repel
```

### {ggiraph}

```{r ggiraph}
gg_point <- ggplot(mtcars, aes(wt, mpg)) +
  ggiraph::geom_point_interactive(tooltip = rownames(mtcars))

girafe(ggobj = gg_point)
```

### {ggbeeswarm}

```{r ggbeeswarm-1}
normal_overplotting <- ggplot(mpg, aes(class, hwy)) +
  geom_point(alpha = 0.4) +
  theme_pubr()
```

```{r ggbeeswarm-2}
ggbeeswarm <- ggplot(mpg, aes(class, hwy)) +
  ggbeeswarm::geom_beeswarm(size = 1.1) +
  theme_pubr()
```

```{r ggbeeswarm-3, fig.width=8, fig.height=5}
library(patchwork)
normal_overplotting + ggbeeswarm
```

### {ggpmisc}

```{r ggpmisc-1}
p <- ggplot(mpg, aes(factor(cyl), hwy)) +
  stat_summary(geom = "col", fun.y = mean, width = 2 / 3, aes(fill = factor(cyl))) +
  labs(x = "Number of cylinders", y = NULL, title = "Means") +
  scale_fill_nejm(guide = FALSE)

data.tb <- tibble(
  x = 7, y = 44,
  plot = list(p +
    theme_pubr(8))
)
```

```{r ggpmisc-2, fig.width=8, fig.height=5}
ggplot(mpg, aes(displ, hwy)) +
  ggpmisc::geom_plot(data = data.tb, aes(x, y, label = plot)) +
  geom_point(aes(colour = factor(cyl))) +
  scale_colour_nejm() +
  labs(
    colour = "Engine cylinders\n(number)"
  ) +
  theme_pubr()
```

### {esquisse}

```{r esquisse, eval = FALSE}
esquisse::esquisser(mpg)
```
