---
title: "Visualization, transformation and reporting with the tidyverse"
author: "Kirill Müller, Tobias Schieferdecker, Patrick Schratz"
date: "`r format(lubridate::with_tz(Sys.time(), 'CET'), '%d %B %Y, %H:%M %Z')`"
documentclass: "book"
site: bookdown::bookdown_site
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE)
options(tibble.print_min = 3, tibble.print_max = 5, max.print = 20)
if (!knitr::is_latex_output()) {
  options(crayon.enabled = TRUE, pillar.bold = TRUE)
  fansi::set_knit_hooks(knitr::knit_hooks)
}
library(here)

this_repo <- function(...) {
  file.path("https://github.com/krlmlr/vistransrep", ...)
}

vistransrep_repo <- function(...) {
  file.path("https://github.com/krlmlr/vistransrep-proj/tree/master", ...)
}
```


# Preface {-}

See the controls at the top of the website for searching, font size, editing, and a link to the PDF version of the material.

## Links {-}

- This website: https://krlmlr.github.io/vistransrep/

- Scripts and installation instructions: `r vistransrep_repo()`
    - Prepared scripts: `r vistransrep_repo("script")`
    - Live code: `r vistransrep_repo("live")`

- The source project for this material: `r this_repo()`

## Package versions used {-}

<details><summary>Click to expand</summary>

```{r include = FALSE}
library(tidyverse)
library(here)
```

```{r}
withr::with_options(list(width = 80), print(sessioninfo::session_info()))
```

</details>

## License {-}

Licensed under [CC-BY-NC 4.0](https://creativecommons.org/licenses/by-nc/4.0/).
