---
title: "Visualization, transformation and reporting with the tidyverse"
author: "Kirill Müller, Tobias Schieferdecker, Patrick Schratz"
date: "`r format(lubridate::with_tz(Sys.time(), 'CET'), '%d %B %Y, %H:%M %Z')`"
documentclass: "book"
site: bookdown::bookdown_site
bibliography: book.bib
biblio-style: apa-old.cls
link-citations: true
nocite: "@grolemund2017"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(fig.width = 3, fig.height = 2, fig.align = "right")
options(tibble.print_min = 3, tibble.print_max = 5, max.print = 20)
if (!knitr::is_latex_output()) {
  options(crayon.enabled = TRUE, pillar.bold = TRUE)
  fansi::set_knit_hooks(knitr::knit_hooks)
} else {
  options(width = 62)
}
library(tidyverse)
library(here)

this_repo <- function(...) {
  file.path("https://github.com/krlmlr/vistransrep", ...)
}

vistransrep_repo <- function(...) {
  file.path("https://github.com/krlmlr/vistransrep-proj/tree/master", ...)
}

# Override
attach(list(
  view = function(x) {
    print(x)
    invisible(x)
  }
))
```


# Preface {-}

See the controls at the top of the website for searching, font size, editing, and a link to the PDF version of the material.

## Links {-}

- This website: https://krlmlr.github.io/vistransrep/book

- Scripts and installation instructions: `r vistransrep_repo()`
    - Prepared scripts: `r vistransrep_repo("script")`

- The source project for this material: `r this_repo()`

## Package versions used {-}

<details><summary>Click to expand</summary>

```{r index-1, include = FALSE}
library(tidyverse)
library(here)
```

```{r index-2}
withr::with_options(list(width = 80), print(sessioninfo::session_info()))
```

</details>

## License {-}

Licensed under [CC-BY-NC 4.0](https://creativecommons.org/licenses/by-nc/4.0/).
